### 2025-01-14 MSc_poster_template.Rmd
Add files via upload

---
title: MSc Health Data Science Project Title
author:
  - name: Athira Mohan
  - name: Brendan Woods
    affil: 1
affiliation:
  - num: 1
    address: School of Mathematical and Statistical Sciences, University of Galway
logoleft_name: https&#58;//sport.universityofgalway.ie/assets/img/logo2.png

poster_height: "594mm"
poster_width: "841mm"
column_numbers: 4
column_margins: "0.5in"	
primary_colour:	"#a80050"	# Secondary colour use for poster design.
secondary_colour:	"#84003d"	# Main colour used for poster design.
accent_colour:	"#e6007e"	# A third colour option.
author_textcol: "white"
title_textsize: "90pt"
author_textsize: "60pt"
body_textsize: "25pt"
output: 
  posterdown::posterdown_html:
    self_contained: true
    pandoc_args: --mathjax
    number_sections: false
knit: pagedown::chrome_print

---

```{css, echo=FALSE}
div.logo_left{
  width: 20%;
}
div.poster_title{
  width: 80%;
}
.section h4 {
    break-after: column;
}
```

<!-- Don't change anything above, except the title and author names, unless you know what you are doing. -->

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 
# Load any additional libraries here
library(tidyverse)
library(plotly)
library(kableExtra)
```

# Background and Problem

This template has been produced using the `posterdown` package. It generates a standard conference-style poster layout.

<!-- You can add references by inserting the following ^[reference details here] -->

A common problem in clinical research is messed up data^[Evans 2010, doi: 10.6030/1939-067x-3.1.1]. It can lead to bias or nonsensical results. Appropriate statistical analysis can result in superb findings in such cases, with the right assumptions.

## Objectives of Project

1. Explain the statistical issue or research question in health data science to be investigated. 
2. Explore a method of analysis.
3. Apply any methods to selected datasets.

# Data Sources and Datasets

The syntax in this poster template and the `posterdown` package uses the same workflow approach as the R Markdown you know and love.

You can even use the bibliography the same way: Our data were taken from a cluster-randomised trial ^[Murphy et al. 2005 doi: 10.1186/1468-6708-6-11], available from the [Irish Social Science Data Archive](https://www.ucd.ie/issda/data/).

<!-- This #### below causes the columns to break -->
<!-- You can change where you put these breaks, and the "column_numbers: 4" argument in the YAMLS header to change the layout of the poster, with care! -->

####

# Early Results / Descriptive Statistics of Datasets

Usually you want to have a nice table displaying some important results that you have calculated. In `posterdown` this is as easy as using the `kable` table formatting you are probably use to as per typical R Markdown formatting.

You can reference tables like so: Table \@ref(tab:mytable). Some basic summaries of the dataset are below:

```{r mytable, out.width='80%'}
knitr::kable(iris[1:10, 1:4], caption = 'Table caption.',align = 'c',"html") %>%
  kable_styling(font_size = 20)
```

Figure \@ref(fig:standard-plot), and Figure \@ref(fig:morefigs) below show the patterns in our dataset. Make sure that all the details in your plots will be legible when printed (legend text, axis text, and any labels)

```{r standard-plot, out.width='80%', fig.align='center', fig.cap='Great figure!', fig.height=4}
ggplot(mpg, aes(displ, hwy, colour = class)) + 
  geom_point()
```

```{r morefigs, out.width='80%', echo=FALSE, fig.cap='Amazing, right?!', fig.height=4}

df1 <- data.frame(dose=c("D0.5", "D1", "D2"),
                len=c(4.2, 10, 29.5))
ggplot(df1, aes(x=dose, y=len, fill=dose)) +
  geom_bar(stat="identity") +
  theme_minimal()

```

You can even make your plots interactive for the HTML version of the poster. You can use the HTML poster for the presentation session, and the PDF poster will be printed - so be sure the static version looks okay.
```{r interactiveplot, out.width='80%', echo=FALSE, fig.cap='Amazing, right?!', fig.height=4}

gapminder <- read_csv("https://raw.githubusercontent.com/datavizpyr/data/master/gapminder-FiveYearData.csv")
df2 <- gapminder %>%
  filter(year %in% c(1952,2007)) %>%
  filter(continent=="Asia")

df2 <- df2 %>%
  mutate(paired = rep(1:(n()/2),each=2),
         year=factor(year))

p <- df2 %>% 
  ggplot(aes(x= lifeExp, y= country)) +
  geom_line(aes(group = paired)) +
  geom_point(aes(color=year), size=4) +
  theme_bw() +
  theme(legend.position="top")

ggplotly(p)

```

####

# Next Project Steps

We plan to conduct further analysis using:

- Variable discombobulation ^[Massey et al. 2005 doi: 15.36.413]
- Expand our minds with explosive machine learning ^[Smith et al. 1991 doi: 12.36.486].

We will use the `plasticanalysis` package for this.

# GitHub

The code and datasets for this project can be viewed at our GitHub repository here: https://github.com/

# References

### 2025-01-18 Power Code.R
Composite Power Code Removed uneccessary code from sample code file

@@ -6,30 +6,41 @@ gw32_t_value<-t.test(tmpDat$gw32[tmpDat$Group == "Metformin"], tmpDat$gw32[tmpDa
 
 
 
-power_function <- function(x) {
-  power_result <- numeric(100000)  
+power_function <- function(x, y, z) {
+  power_result <- numeric(1000)  
   i <- 1
   repeat {
-placebo_group <- thesis.df[thesis.df$Group == "Placebo", ]
-clean_placebo_group<-as.vector(na.omit(placebo_group[[x]]))
-sampled_placebo <- sample(clean_placebo_group, 15, replace = T)
+placebo_group <- tmpDat[tmpDat$Group == "Placebo", ]
+clean_placebo_group1<-as.vector(na.omit(placebo_group[[x]]))
+clean_placebo_group2<-as.vector(na.omit(placebo_group[[y]]))
+clean_placebo_group3<-as.vector(na.omit(placebo_group[[z]]))
+sampled_placebo1 <- sample(clean_placebo_group1, 15, replace = T)
+sampled_placebo2 <- sample(clean_placebo_group2, 15, replace = T)
+sampled_placebo3 <- sample(clean_placebo_group3, 15, replace = T)
 
-metformin_group <- thesis.df[thesis.df$Group == "Metformin", ]
-clean_metformin_group<-as.vector(na.omit(metformin_group[[x]]))
-sampled_metformin <- sample(clean_metformin_group, 15, replace = T)
+metformin_group <- tmpDat[tmpDat$Group == "Metformin", ]
+clean_metformin_group1<-as.vector(na.omit(metformin_group[[x]]))
+clean_metformin_group2<-as.vector(na.omit(metformin_group[[y]]))
+clean_metformin_group3<-as.vector(na.omit(metformin_group[[z]]))
 
-t_test_result<- t.test(sampled_metformin, sampled_placebo)
-power_result[i] <- t_test_result$statistic
+sampled_metformin1 <- sample(clean_metformin_group1, 15, replace = T)
+sampled_metformin2 <- sample(clean_metformin_group2, 15, replace = T)
+sampled_metformin3 <- sample(clean_metformin_group3, 15, replace = T)
+
+t_test_result1<- t.test(sampled_metformin1, sampled_placebo1)
+t_test_result2<- t.test(sampled_metformin2, sampled_placebo2)
+t_test_result3<- t.test(sampled_metformin3, sampled_placebo3)
+
+
[…]

### 2025-01-18 Sample Code.R

@@ -4,9 +4,10 @@ library(dplyr)
 load("\\Users\\brend\\Downloads\\data")
 ls
 View(tmpDat)
+
 ## Creating binary column for Insulin ##
 tmpDat$binary_insulin<-0
-tmpDat$binary_insulin[thesis.df$InsulinYN == "Yes"]<-1
+tmpDat$binary_insulin[tmpDat$InsulinYN == "Yes"]<-1
 
 thesis.df<-tmpDat
 
@@ -26,29 +27,9 @@ study_squared_composite<-z_score_squared_insulin+(gw32_t_value^2)+(gw38_t_value^
 ##Not squared##
 study_non_squared_composite<-sqrt(z_score_squared_insulin)+gw32_t_value+gw38_t_value
 
-##Permutation Code for t-values for continuous data (gw32, gw38) and for z-score for binary data (InsulinYN) ##
+##Permutation Code for t-values for continuous data (gw32, gw38, binary_insulin) ##
 
-## Z-score for Insulin ##
-proportion_permutation_function<-function(x){
-  permutation_chisquare<-numeric(100000)
-  i <-1
-  repeat{
-    thesis.df$Group<-sample(thesis.df$Group, 535, replace = F)
-    permutation_result<-prop.test(x = c(sum(x[thesis.df$Group== "Metformin"] == "Yes", na.rm=T),
-                                        sum(x[thesis.df$Group== "Placebo"] == "Yes", na.rm=T)),
-                                  n = c(sum(thesis.df$Group == "Metformin"),sum(thesis.df$Group == "Placebo"))) 
-    permutation_chisquare[i]<-permutation_result$statistic
-    i<- i +1
-    if(i > 100000) break}
-  return(permutation_chisquare)
-}
 
-summary(proportion_permutation_function(thesis.df$InsulinYN))
-hist(proportion_permutation_function(thesis.df$InsulinYN),
-     breaks = 500, main= "Histogram of Permuted Z-scores of Insulin comparing Treatment Groups",
-     xlab= "Z-scores Insulin")
-
-##T-values for gw32, gw 38 ##
 continuous_permutation_function<-function(x){
   permutation_t_value<-numeric(100000)
   i <-1
@@ -72,7 +53,7 @@ hist(continuous_permutation_function(thesis.df$gw38),
      xlab= "T-values gw38")
 
 ## Creating the Composite ##
-Zi+t32+t38
+
 ## The squared method ##
 
[…]

### 2025-01-18 Sample Code.R
Update Sample Code.R
Questions on power code

@@ -139,4 +139,4 @@ power_function <- function(x, y, z) {
 
 power_function("gw32", "gw38", "binary_insulin")
 
-## Not very tidy, is this an issue? Also, is there a problem with how I used the squared composite when getting the proportion?
+## Not very tidy, is this an issue? Also, is there a problem with how I used the squared composite when getting the proportion? Should the size to be sampled vary?

### 2025-01-18 Sample Code.R
Graphs for poster for power and composite permutation

@@ -4,7 +4,8 @@ library(dplyr)
 load("\\Users\\brend\\Downloads\\data")
 ls
 View(tmpDat)
-
+library(ggplot2)
+library(hrbrthemes)
 ## Creating binary column for Insulin ##
 tmpDat$binary_insulin<-0
 tmpDat$binary_insulin[tmpDat$InsulinYN == "Yes"]<-1
@@ -31,14 +32,14 @@ study_non_squared_composite<-sqrt(z_score_squared_insulin)+gw32_t_value+gw38_t_v
 
 
 continuous_permutation_function<-function(x){
-  permutation_t_value<-numeric(100000)
+  permutation_t_value<-numeric(500000)
   i <-1
   repeat{
     thesis.df$Group<-sample(thesis.df$Group, 535, replace = F)
     permutation_result<-t.test(x[thesis.df$Group == "Metformin"], x[thesis.df$Group == "Placebo"])
     permutation_t_value[i]<-permutation_result$statistic
     i<- i +1
-    if(i > 100000) break}
+    if(i > 500000) break}
   return(permutation_t_value)
 }
 
@@ -65,6 +66,24 @@ non_squared_composite<-continuous_permutation_function(thesis.df$binary_insulin)
   continuous_permutation_function(thesis.df$gw32) +
   continuous_permutation_function(thesis.df$gw38)
 
+##Permuted Squared Test Statistic Graph##
+
+percentile_95<-quantile(squared_composite, 0.95)
+
+squared_composite.df<-data.frame(squared_composite)
+ggplot(squared_composite.df, aes(x= squared_composite))+
+  geom_histogram(bins = 200, fill = "darkslategray", color = "darkslategray")+
+  scale_x_continuous(limits = c(0, 20)) +
+  labs(x = "Squared Composite", y = "Frequency", title = "Squared Composite Test Statistic") +
+  geom_vline(xintercept = percentile_95, color = "red", linetype = "dashed", size = 0.75)+
+  annotate("text", x = percentile_95, y = 10000, label = "95th Percentile", color = "red", angle = 90, vjust = -0.5)+
+  theme_minimal()
+
+
+
+  
+
+
 ##P-value calculation##
 
 ##Individual variables##
@@ -80,6 +99,9 @@ p_value_squared_composite<-sum(squared_composite>study_squared_composite)/
   length(squared_composite)
 p_value_non_squared_composite<-sum(non_squared_composite<study_non_squared_composite)/
   length(non_squared_composite)
+
+
[…]

### 2025-01-20 Sample Code.R
Trying to resolve conflict 2

@@ -14,8 +14,8 @@ thesis.df<-tmpDat
 
 ## T-values and Z-scores for the original study ##
 StudyInsulinTest<- prop.test(x = c(sum(tmpDat$InsulinYN[tmpDat$Group== "Metformin"] == "Yes", na.rm=T),
-                              sum(tmpDat$InsulinYN[tmpDat$Group== "Placebo"] == "Yes", na.rm=T)),
-                        n = c(sum(tmpDat$Group == "Metformin"),sum(tmpDat$Group == "Placebo")))
+                                   sum(tmpDat$InsulinYN[tmpDat$Group== "Placebo"] == "Yes", na.rm=T)),
+                             n = c(sum(tmpDat$Group == "Metformin"),sum(tmpDat$Group == "Placebo")))
 z_score_squared_insulin<-sqrt(StudyInsulinTest$statistic)       ## This is not used, binary insulin more appropriate to use ##                       
 
 insulin_t_value<-t.test(tmpDat$binary_insulin[tmpDat$Group == "Metformin"], tmpDat$binary_insulin[tmpDat$Group == "Placebo"])$statistic
@@ -81,7 +81,7 @@ ggplot(squared_composite.df, aes(x= squared_composite))+
 
 
 
-  
+
 
 
 ##P-value calculation##

### 2025-01-20 Sample Code.R
Updating graph for permutations

@@ -73,11 +73,13 @@ percentile_95<-quantile(squared_composite, 0.95)
 squared_composite.df<-data.frame(squared_composite)
 ggplot(squared_composite.df, aes(x= squared_composite))+
   geom_histogram(bins = 200, fill = "darkslategray", color = "darkslategray")+
-  scale_x_continuous(limits = c(0, 20)) +
+  scale_x_continuous(limits = c(0, 25)) +
   labs(x = "Squared Composite", y = "Frequency", title = "Squared Composite Test Statistic") +
   geom_vline(xintercept = percentile_95, color = "red", linetype = "dashed", size = 0.75)+
   annotate("text", x = percentile_95, y = 10000, label = "95th Percentile", color = "red", angle = 90, vjust = -0.5)+
-  theme_minimal()
+  theme_minimal()+
+  geom_vline( xintercept = study_squared_composite, color = "gold", size = 1 )+
+  annotate("text", x = study_squared_composite, y = 10000, label = "EMERGE Composite", color = "gold", angle = 90, vjust = -0.5)
 
 
 
@@ -180,7 +182,7 @@ sample_size<-c(30, 40, 50, 60, 70, 80)
 power_graph.df<-data.frame(sample_size, power_results)
 ggplot(power_graph.df, aes(x=sample_size, y=power_results)) +
   geom_line( color="mediumseagreen", size=1.5, alpha=0.9, linetype=1) +
-  theme_classic() + labs(title = "Power of Permuted Test Statistics Method", x = "Sample Size", y= "Power")+
+  theme_bw() + labs(title = "Power of Permuted Test Statistics Method", x = "Sample Size", y= "Power")+
   scale_x_continuous(limits = c(30, NA), expand = c(0, 0))  +
   ylim(0, NA) 
 

### 2025-01-20 MSc_poster_template.Rmd
New Template

@@
 knit: pagedown::chrome_print
-
 ---
@@
     break-after: column;
+}
+div.footnotes {
+    font-size: 18pt;
 }
@@
 library(kableExtra)
+library(posterdown)
+load("\\Users\\brend\\Downloads\\data")
+tmpDat$binary_insulin<-0
+tmpDat$binary_insulin[tmpDat$InsulinYN == "Yes"]<-1
+
+thesis.df<-tmpDat
 ```
@@
 
-A common problem in clinical research is messed up data^[Evans 2010, doi: 10.6030/1939-067x-3.1.1].
+A common problem in clinical research is messed up data[^1].
 It can lead to bias or nonsensical results.
 Appropriate statistical analysis can result in superb findings in such cases, with the right assumptions.
+
+[^1]: Evans 2010, doi: 10.6030/1939-067x-3.1.1
 
@@
 
-You can even use the bibliography the same way: Our data were taken from a cluster-randomised trial ^[Murphy et al.
-2005 doi: 10.1186/1468-6708-6-11], available from the [Irish Social Science Data Archive](https://www.ucd.ie/issda/data/).
+You can even use the bibliography the same way: Our data were taken from a cluster-randomised trial [^2], available from the [Irish Social Science Data Archive](https://www.ucd.ie/issda/data/).
+
+[^2]: Murphy et al. 2005 doi: 10.1186/1468-6708-6-11
 
 <!-- This #### below causes the columns to break -->
+
 <!-- You can change where you put these breaks, and the "column_numbers: 4" argument in the YAMLS header to change the layout of the poster, with care! -->
@@
 ```{r morefigs, out.width='80%', echo=FALSE, fig.cap='Amazing, right?!', fig.height=4}
-
-df1 <- data.frame(dose=c("D0.5", "D1", "D2"),
-                len=c(4.2, 10, 29.5))
-ggplot(df1, aes(x=dose, y=len, fill=dose)) +
-  geom_bar(stat="identity") +
-  theme_minimal()
+continuous_permutation_function<-function(x){
+  permutation_t_value<-numeric(100000)
+  i <-1
+  repeat{
+    thesis.df$Group<-sample(thesis.df$Group, 535, replace = F)
+    permutation_result<-t.test(x[thesis.df$Group == "Metformin"], x[thesis.df$Group == "Placebo"])
+    permutation_t_value[i]<-permutation_result$statistic
+    i<- i +1
+    if(i > 100000) break}
+  return(permutation_t_value)
+}
+head(thesis.df)
 
@@
 You can use the HTML poster for the presentation session, and the PDF poster will be printed - so be sure the static version looks okay.
+
 ```{r interactiveplot, out.width='80%', echo=FALSE, fig.cap='Amazing, right?!', fig.height=4}
@@
 
-- Variable discombobulation ^[Massey et al. 2005 doi: 15.36.413]
-- Expand our minds with explosive machine learning ^[Smith et al. 1991 doi: 12.36.486].
+-   Variable discombobulation [^3]
+-   Expand our minds with explosive machine learning [^4].
+
+[^3]: Massey et al. 2005 doi: 15.36.413
+
+[^4]: Smith et al. 1991 doi: 12.36.486
 
@@
 
-The code and datasets for this project can be viewed at our GitHub repository here: https://github.com/
+The code and datasets for this project can be viewed at our GitHub repository here: <https://github.com/>
 

### 2025-01-22 Sample Code.R
Updated to use the composite to use the t-value rather than z-score for the insulin element.

@@ -22,11 +22,11 @@ insulin_t_value<-t.test(tmpDat$binary_insulin[tmpDat$Group == "Metformin"], tmpD
 gw32_t_value<-t.test(tmpDat$gw32[tmpDat$Group == "Metformin"], tmpDat$gw32[tmpDat$Group == "Placebo"])$statistic
 gw38_t_value<-t.test(tmpDat$gw38[tmpDat$Group == "Metformin"], tmpDat$gw38[tmpDat$Group == "Placebo"])$statistic
 
-##Composites for Original study##
+##Composites for Original Data##
 ##Squared##
-study_squared_composite<-z_score_squared_insulin+(gw32_t_value^2)+(gw38_t_value^2)
+study_squared_composite<-(insulin_t_value^2)+(gw32_t_value^2)+(gw38_t_value^2)
 ##Not squared##
-study_non_squared_composite<-sqrt(z_score_squared_insulin)+gw32_t_value+gw38_t_value
+study_non_squared_composite<-insulin_t_value+gw32_t_value+gw38_t_value
 
 ##Permutation Code for t-values for continuous data (gw32, gw38, binary_insulin) ##
 
@@ -102,6 +102,16 @@ p_value_squared_composite<-sum(squared_composite>study_squared_composite)/
 p_value_non_squared_composite<-sum(non_squared_composite<study_non_squared_composite)/
   length(non_squared_composite)
 
+##P-values for Poster ##
+#Need to calculate P-value for discretised gw32, gw38, as these are not individually reported in EMERGE#
+EMERGE_gw32_p_value<-prop.test(x = c(sum(tmpDat$gw32greater5.1[tmpDat$Group == "Metformin"] == "Yes", na.rm = T),
+                               sum(tmpDat$gw32greater5.1[tmpDat$Group== "Placebo"]== "Yes", na.rm = T)),
+                               n = c(sum(tmpDat$Group == "Metformin"), sum(tmpDat$Group == "Placebo")))
+
+EMERGE_gw38_p_value<-prop.test(x = c(sum(tmpDat$gw38greater5.1[tmpDat$Group == "Metformin"] == "Yes", na.rm = T),
+                                     sum(tmpDat$gw38greater5.1[tmpDat$Group== "Placebo"]== "Yes", na.rm = T)),
+                               n = c(sum(tmpDat$Group == "Metformin"), sum(tmpDat$Group == "Placebo")))
+
 
[…]

### 2025-01-22 MSc_poster_template.Rmd
First look, with methodologies, early results, and future direction

@@
 ---
-title: MSc Health Data Science Project Title
+title: Meta-analysis of composite outcomes in clinical trials 
 author:
@@
 author_textcol: "white"
-title_textsize: "90pt"
+title_textsize: "80pt"
 author_textsize: "60pt"
@@
 library(posterdown)
+library(ggplot2)
 load("\\Users\\brend\\Downloads\\data")
@@
 thesis.df<-tmpDat
+
+insulin_t_value<-t.test(tmpDat$binary_insulin[tmpDat$Group == "Metformin"], tmpDat$binary_insulin[tmpDat$Group == "Placebo"])$statistic
+gw32_t_value<-t.test(tmpDat$gw32[tmpDat$Group == "Metformin"], tmpDat$gw32[tmpDat$Group == "Placebo"])$statistic
+gw38_t_value<-t.test(tmpDat$gw38[tmpDat$Group == "Metformin"], tmpDat$gw38[tmpDat$Group == "Placebo"])$statistic
+study_squared_composite<-(insulin_t_value^2)+(gw32_t_value^2)+(gw38_t_value^2)
+
 ```
@@
 
-# Early Results / Descriptive Statistics of Datasets
+# Descriptive Statistics of Datasets
 
@@
 
-```{r morefigs, out.width='80%', echo=FALSE, fig.cap='Amazing, right?!', fig.height=4}
+# Methodology
+The meta-analysis of the EMERGE outcomes used permutation tests, which generate a null distribution by randomly shuffling the treatment group assignments of participants and their outcome variables.
+This breaks any true relationship between treatment and outcomes.
+Statistical significance is assessed by comparing the actual test statistic to the distribution of test statistics from these shuffled datasets.
+The P-value is calculated as the proportion of the null distribution that is more extreme than the observed test statistic.
+
+The power of the permutation test is estimated using sampling with replacement.
+Participants are randomly resampled from the original data, allowing duplicates, while maintaining equal numbers from both treatment groups.
+The test is repeated multiple times, and power is determined by the proportion of resampled subsets yielding test statistics more extreme than the significance threshold.
+
+# Early Results 
+```{r morefigs, out.width='90%', echo=FALSE, fig.cap='Permuted Null Distribution of Composite', fig.height=6}
 continuous_permutation_function<-function(x){
-  permutation_t_value<-numeric(100000)
+  permutation_t_value<-numeric(10000)
   i <-1
@@
     i<- i +1
-    if(i > 100000) break}
+    if(i > 10000) break}
   return(permutation_t_value)
 }
-head(thesis.df)
+squared_composite<-((continuous_permutation_function(thesis.df$binary_insulin))^2) +
+  ((continuous_permutation_function(thesis.df$gw32))^2) +
+  ((continuous_permutation_function(thesis.df$gw38))^2)
+percentile_95<-quantile(squared_composite, 0.95)
+
+squared_composite.df<-data.frame(squared_composite)
+ggplot(squared_composite.df, aes(x= squared_composite))+
+  geom_histogram(bins = 200, fill = "darkslategrey", color = "darkslategrey")+
+  scale_x_continuous(limits = c(0, 30)) +
+  labs(x = "Squared Composite", y = "Frequency", title = "Permuted Test Statistics") +
+  geom_vline(xintercept = percentile_95, color = "darkred", linetype = "dashed", size = 1)+
+  annotate("text", x = percentile_95, y = 250, label = "95th Percentile", color = "darkred", angle = 90, vjust = -0.5, size = 7)+
+  theme_minimal()+
+  theme(axis.title = element_text( size = 24),
+        axis.text = element_text(size = 18),
+        plot.title = element_text(size = 28))+
+  geom_vline( xintercept = study_squared_composite, color = "gold3", size = 1 )+
+  annotate("text", x = study_squared_composite, y = 250, label = "EMERGE Composite", color = "gold3", angle = 90, vjust = -0.5, size = 7)
 
@@
 
-You can even make your plots interactive for the HTML version of the poster.
-You can use the HTML poster for the presentation session, and the PDF poster will be printed - so be sure the static version looks okay.
+```{r morefigs2, out.width='90%', echo=FALSE, fig.cap='Power of Permutation Test', fig.height=5}
+APPLE<-7.881899
+power_n_30<-0.219598
+power_n_40<-0.264418
+power_n_50<-0.314862
+power_n_60<-0.367066
+power_n_70<-0.418002
+power_n_80<-0.468156
+power_results<-c(power_n_30, power_n_40, power_n_50, power_n_60, power_n_70, power_n_80)
+sample_size<-c(30, 40, 50, 60, 70, 80)
 
-```{r interactiveplot, out.width='80%', echo=FALSE, fig.cap='Amazing, right?!', fig.height=4}
-
-gapminder <- read_csv("https://raw.githubusercontent.com/datavizpyr/data/master/gapminder-FiveYearData.csv")
-df2 <- gapminder %>%
-  filter(year %in% c(1952,2007)) %>%
-  filter(continent=="Asia")
-
-df2 <- df2 %>%
-  mutate(paired = rep(1:(n()/2),each=2),
-         year=factor(year))
-
-p <- df2 %>% 
-  ggplot(aes(x= lifeExp, y= country)) +
-  geom_line(aes(group = paired)) +
-  geom_point(aes(color=year), size=4) +
-  theme_bw() +
-  theme(legend.position="top")
-
-ggplotly(p)
+power_graph.df<-data.frame(sample_size, power_results)
+ggplot(power_graph.df, aes(x=sample_size, y=power_results)) +
+  geom_line( color="mediumseagreen", size=1.5, alpha=0.9, linetype=1) +
+  theme_bw() + labs(title = "Power vs. Sample Size", x = "Sample Size", y= "Power")+
+  scale_x_continuous(limits = c(30, NA), expand = c(0, 0))  +
+  ylim(0, NA)+
+  theme(axis.title = element_text(size = 24),
+         axis.text = element_text(size = 18),
+        plot.title = element_text(size = 28))
 
@@
 
-####
-
 # Next Project Steps
 
-We plan to conduct further analysis using:
+The next steps in our project is to compare the permutation method with other
+approaches to improving the analysis of composite outputs, in particular to the
+methods described by Mc Menamin *et al,* [^5].
 
--   Variable discombobulation [^3]
--   Expand our minds with explosive machine learning [^4].
-
-[^3]: Massey et al. 2005 doi: 15.36.413
-
-[^4]: Smith et al. 1991 doi: 12.36.486
-
-We will use the `plasticanalysis` package for this.
+[^5]: Mc Menamin *et al,* 2018 doi:https://doi.org/10.1186/s13023-018-0819-1
 
@@
 
-The code and datasets for this project can be viewed at our GitHub repository here: <https://github.com/>
+The code and datasets for this project can be viewed at our GitHub repository here: <https://github.com/Brendan-G-Woods/Thesis>
 

### 2025-01-27 MSc_poster_template.Rmd
Power calculation of EMERGE composite. Update to graphs

@@
 # Early Results 
-```{r morefigs, out.width='90%', echo=FALSE, fig.cap='Permuted Null Distribution of Composite', fig.height=6}
+```{r morefigs, out.width='90%', echo=FALSE, fig.cap='Permuted Null Distribution of Composite. Test statistic 28.89, P-value of P<0.001', fig.height=6}
 continuous_permutation_function<-function(x){
@@
   scale_x_continuous(limits = c(0, 30)) +
-  labs(x = "Squared Composite", y = "Frequency", title = "Permuted Test Statistics") +
-  geom_vline(xintercept = percentile_95, color = "darkred", linetype = "dashed", size = 1)+
-  annotate("text", x = percentile_95, y = 250, label = "95th Percentile", color = "darkred", angle = 90, vjust = -0.5, size = 7)+
+  labs(x = "Squared Composite", y = "Frequency", title = expression(bold("Permuted Null Distribution"))) +
+  geom_vline(xintercept = percentile_95, color = "maroon", linetype = "dashed", size = 1)+
+  annotate("text", x = percentile_95, y = 250, label = "95th Percentile", color = "maroon", angle = 90, vjust = -0.5, size = 7)+
   theme_minimal()+
@@
 
-```{r morefigs2, out.width='90%', echo=FALSE, fig.cap='Power of Permutation Test', fig.height=5}
-APPLE<-7.881899
+```{r morefigs2, out.width='95%', echo=FALSE, fig.cap='Power of Permutation Test', fig.height=5}
+
 power_n_30<-0.219598
@@
 power_n_80<-0.468156
-power_results<-c(power_n_30, power_n_40, power_n_50, power_n_60, power_n_70, power_n_80)
-sample_size<-c(30, 40, 50, 60, 70, 80)
+power_n_90<-0.51759
+power_n_100<-0.56405
+power_n_110<-0.607478
+power_n_120<-0.648076
+power_n_130<-0.685648
+power_n_140<-0.723104
+power_n_150<-0.754432
+power_n_160<-0.781646
+power_n_170<-0.809294
+power_n_180<-0.834104
+power_n_190<-0.854704
+power_n_200<-0.87419
+power_n_210<-0.890702
+power_n_220<-0.905218
+power_n_230<-0.918456
+power_n_240<-0.92976
+power_n_250<-0.939556
+power_n_260<-0.948088
+power_results<-c(power_n_30, power_n_40, power_n_50, power_n_60, power_n_70, power_n_80, power_n_90,
+                 power_n_100, power_n_110, power_n_120, power_n_130, power_n_140, power_n_150, power_n_160,
+                 power_n_170, power_n_180, power_n_190, power_n_200, power_n_210, power_n_220, power_n_230,
+                 power_n_240, power_n_250, power_n_260)
 
-power_graph.df<-data.frame(sample_size, power_results)
-ggplot(power_graph.df, aes(x=sample_size, y=power_results)) +
-  geom_line( color="mediumseagreen", size=1.5, alpha=0.9, linetype=1) +
-  theme_bw() + labs(title = "Power vs. Sample Size", x = "Sample Size", y= "Power")+
+EMERGE_power_n_30<-0.02331
+EMERGE_power_n_40<-0.03202
+EMERGE_power_n_50<-0.05006
+EMERGE_power_n_60<-0.049188
+EMERGE_power_n_70<-0.056814
+EMERGE_power_n_80<-0.06352
+EMERGE_power_n_90<-0.06244
+EMERGE_power_n_100<-0.07752
+EMERGE_power_n_110<-0.07844
+EMERGE_power_n_120<-0.08728
+EMERGE_power_n_130<-0.09256
+EMERGE_power_n_140<-0.09306
+EMERGE_power_n_150<-0.11259
+EMERGE_power_n_160<-0.10677
+EMERGE_power_n_170<-0.11652
+EMERGE_power_n_180<-0.12682
+EMERGE_power_n_190<-0.12433
+EMERGE_power_n_200<-0.13877
+EMERGE_power_n_210<-0.14522
+EMERGE_power_n_220<-0.14281
+EMERGE_power_n_230<-0.1552
+EMERGE_power_n_240<-0.16422
+EMERGE_power_n_250<-0.16039
+EMERGE_power_n_260<-0.17103
+
+EMERGE_power_results<-c(EMERGE_power_n_30, EMERGE_power_n_40, EMERGE_power_n_50, EMERGE_power_n_60,
+                        EMERGE_power_n_70, EMERGE_power_n_80, EMERGE_power_n_90, EMERGE_power_n_100,
+                        EMERGE_power_n_110, EMERGE_power_n_120, EMERGE_power_n_130, EMERGE_power_n_140,
+                        EMERGE_power_n_150, EMERGE_power_n_160, EMERGE_power_n_170, EMERGE_power_n_180,
+                        EMERGE_power_n_190, EMERGE_power_n_200, EMERGE_power_n_210, EMERGE_power_n_220,
+                        EMERGE_power_n_230, EMERGE_power_n_240, EMERGE_power_n_250, EMERGE_power_n_260)
+sample_size<-c(30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140, 150, 160, 170, 180, 190, 200, 210,
+               220, 230, 240, 250, 260)
+
+power_graph.df<-data.frame(sample_size, power_results, EMERGE_power_results)
+
+ggplot(power_graph.df, aes(x = sample_size)) +
+  geom_line(aes(y = power_results, color = "Permutation Test Method"), size = 1.5, alpha = 0.9, linetype = 1) +
+  geom_line(aes(y = EMERGE_power_results, color = "EMERGE Composite   "), size = 1.5, alpha = 0.9, linetype = 1) +
+  labs(title = expression(bold("Power Comparison")), x = "Sample Size", y = "Power") +
   scale_x_continuous(limits = c(30, NA), expand = c(0, 0)) +
-  ylim(0, NA)+
-  theme(axis.title = element_text(size = 24),
+  scale_y_continuous(limits = c(0, 1)) +
+  theme_bw()+
+  theme(
+    axis.title = element_text(size = 24),
     axis.text = element_text(size = 18),
-        plot.title = element_text(size = 28))
+    plot.title = element_text(size = 28),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 18),
+    legend.position = "bottom")+
+  scale_color_manual(
+    name = "Legend", 
+    values = c("Permutation Test Method" = "mediumseagreen", "EMERGE Composite   " = "steelblue")
+  )
 

### 2025-01-27 Sample Code.R

@@ -140,7 +140,7 @@ power_function("binary_insulin")
 
 ##Power Code for Squared Composite##
 power_function <- function(x, y, z) {
-  power_result <- numeric(1000)  
+  power_result <- numeric(500000)  
   i <- 1
   repeat {
     placebo_group <- tmpDat[tmpDat$Group == "Placebo", ]
@@ -167,14 +167,12 @@ power_function <- function(x, y, z) {
     
     power_result[i] <- (((t_test_result1$statistic)^2)+((t_test_result2$statistic)^2)+((t_test_result3$statistic)^2))
     i<- i +1
-    if(i > 1000) break}
-  return(sum((power_result>quantile(squared_composite, 0.95))/ 1000))
+    if(i > 500000) break}
+  return(sum((power_result>quantile(squared_composite, 0.95))/ 500000))
 }
 
 power_function("gw32", "gw38", "binary_insulin")
 
-
-## Not very tidy, is this an issue? Also, is there a problem with how I used the squared composite when getting the proportion? Should it be a variable of the function?
 
 ## Power Chart, generated with 500,000 test statistic permutations, starting at a sample size of 30, with increments of 10
 ## and ending at 80, with 500,000 resamples ##
@@ -186,13 +184,102 @@ power_n_50<-0.314862
 power_n_60<-0.367066
 power_n_70<-0.418002
 power_n_80<-0.468156
-power_results<-c(power_n_30, power_n_40, power_n_50, power_n_60, power_n_70, power_n_80)
-sample_size<-c(30, 40, 50, 60, 70, 80)
+power_n_90<-0.51759
+power_n_100<-0.56405
+power_n_110<-0.607478
+power_n_120<-0.648076
+power_n_130<-0.685648
+power_n_140<-0.723104
+power_n_150<-0.754432
+power_n_160<-0.781646
+power_n_170<-0.809294
+power_n_180<-0.834104
+power_n_190<-0.854704
+power_n_200<-0.87419
+power_n_210<-0.890702
+power_n_220<-0.905218
+power_n_230<-0.918456
+power_n_240<-0.92976
+power_n_250<-0.939556
+power_n_260<-0.948088
+power_results<-c(power_n_30, power_n_40, power_n_50, power_n_60, power_n_70, power_n_80, power_n_90,
[…]

### 2025-01-28 MSc_poster_template.Rmd
Comments addressed

@@
 # Methodology
-The meta-analysis of the EMERGE outcomes used permutation tests, which generate a null distribution by randomly shuffling the treatment group assignments of participants and their outcome variables.
+The meta-analysis of the EMERGE outcomes used permutation tests, which generate a null distribution by randomly shuffling the treatment group assignment of participants, while maintaining the outcome variables association with each other.
 This breaks any true relationship between treatment and outcomes.
-Statistical significance is assessed by comparing the actual test statistic to the distribution of test statistics from these shuffled datasets.
+Statistical significance is assessed by comparing the actual test statistic to the null distribution from these shuffled datasets.
 The P-value is calculated as the proportion of the null distribution that is more extreme than the observed test statistic.
+The composite test statistic is obtained by summing the squared T-values comparing glucose at weeks 32 and 38, and Insulin: ***T~32~^2^ + T~38~^2^ + T~I~^2^ ***
 
@@
 # Early Results 
-```{r morefigs, out.width='90%', echo=FALSE, fig.cap='Permuted Null Distribution of Composite. Test statistic 28.89, P-value of P<0.001', fig.height=6}
+```{r morefigs, out.width='90%', echo=FALSE, fig.cap='Permuted Null Distribution of Composite. Test statistic 28.89, P-value of P<0.001', fig.height=4.5}
 continuous_permutation_function<-function(x){
@@
 ggplot(squared_composite.df, aes(x= squared_composite))+
-  geom_histogram(bins = 200, fill = "darkslategrey", color = "darkslategrey")+
+  geom_histogram(bins = 200, fill = "salmon", color = "salmon")+
   scale_x_continuous(limits = c(0, 30)) +
@@
   geom_vline(xintercept = percentile_95, color = "maroon", linetype = "dashed", size = 1)+
-  annotate("text", x = percentile_95, y = 250, label = "95th Percentile", color = "maroon", angle = 90, vjust = -0.5, size = 7)+
+  annotate("text", x = percentile_95, y = 200, label = "95th Percentile", color = "maroon", angle = 90, vjust = -0.5, size = 7)+
   theme_minimal()+
@@
   geom_vline( xintercept = study_squared_composite, color = "gold3", size = 1 )+
-  annotate("text", x = study_squared_composite, y = 250, label = "EMERGE Composite", color = "gold3", angle = 90, vjust = -0.5, size = 7)
+  annotate("text", x = study_squared_composite, y = 200, label = "EMERGE Composite", color = "gold3", angle = 90, vjust = -0.5, size = 7)
 
@@
 
-```{r morefigs2, out.width='95%', echo=FALSE, fig.cap='Power of Permutation Test', fig.height=5}
+```{r morefigs2, out.width='95%', echo=FALSE, fig.cap='Power of Permutation Test', fig.height=4.5}
 
@@
 
-The next steps in our project is to compare the permutation method with other
-approaches to improving the analysis of composite outputs, in particular to the
-methods described by Mc Menamin *et al,* [^5].
+The next steps in our project are to:
+
+* Compare the permutation test method with the approach Mc Menamin *et al.*, [^5] used to analyze composite outputs.
+* Compare different methods of creating the composite and their influence on power.
+* Derive a theoretical distribution incorporating correlation between the variables.
 
@@
 
-The code and datasets for this project can be viewed at our GitHub repository here: <https://github.com/Brendan-G-Woods/Thesis>
+Our code and datasets can be found here: <https://github.com/Brendan-G-Woods/Thesis>
 

### 2025-01-28 MSc_poster_template.Rmd
More feedback addressed

@@
 
-You can even use the bibliography the same way: Our data were taken from a cluster-randomised trial [^2], available from the [Irish Social Science Data Archive](https://www.ucd.ie/issda/data/).
-
-[^2]: Murphy et al. 2005 doi: 10.1186/1468-6708-6-11
+You can even use the bibliography the same way: Our data were taken from a cluster-randomised trial, available from the [Irish Social Science Data Archive](https://www.ucd.ie/issda/data/).
 
@@
 Participants are randomly resampled from the original data, allowing duplicates, while maintaining equal numbers from both treatment groups.
-The test is repeated multiple times, and power is determined by the proportion of resampled subsets yielding test statistics more extreme than the significance threshold.
+The test is repeated multiple times, and power is determined by the proportion of resampled subsets yielding test statistics more extreme than the 5% significance threshold.
 
@@
 * Compare the permutation test method with the approach Mc Menamin *et al.*, [^5] used to analyze composite outputs.
-* Compare different methods of creating the composite and their influence on power.
-* Derive a theoretical distribution incorporating correlation between the variables.
+* Compare different methods of creating the composite test statistic and their influence on power.
+* Derive a theoretical null distribution for the composite test statistic incorporating the correlation between the variables.
 

### 2025-01-28 MSc_poster_template.Rmd
Background and Problem updated

@@
 # Background and Problem
+Composites are commonly used in clinical trials as the primary outcome.
+When misused, they can reduce the power to detect treatment effect.
+This was seen in EMERGE, a randomised clinical trial investigating the effects of early initiation of Metformin treatment on developing gestational diabetes.
+The composite was created using three measures of glycemic control; blood glucose levels of 5.1mmol/L or greater at gestation weeks 32 or 38, or initiation of Insulin treatment.
 
-This template has been produced using the `posterdown` package. It generates a standard conference-style poster layout.
+ ***Insulin~Yes~ OR gw32≥5.1~Yes~ OR gw38≥5.1~Yes~ ***
 
-<!-- You can add references by inserting the following ^[reference details here] -->
+This composite involves dichotomising blood glucose levels, greatly reducing power. Therefore, the objectives of this project are:
 
-A common problem in clinical research is messed up data[^1].
-It can lead to bias or nonsensical results.
-Appropriate statistical analysis can result in superb findings in such cases, with the right assumptions.
+1.	To perform a meta-analysis of the primary outcome using permutations of the composite test statistic, and analyzing the glucose components as continuous variables
 
-[^1]: Evans 2010, doi: 10.6030/1939-067x-3.1.1
+2.	To compare various ways of creating the composite test statistic, and their power
 
-## Objectives of Project
-
-1.  Explain the statistical issue or research question in health data science to be investigated.
-2.  Explore a method of analysis.
-3.  Apply any methods to selected datasets.
+3.	To compare the permutation method with other approaches to analyzing composite outcomes
 
@@
 The syntax in this poster template and the `posterdown` package uses the same workflow approach as the R Markdown you know and love.
-
-You can even use the bibliography the same way: Our data were taken from a cluster-randomised trial, available from the [Irish Social Science Data Archive](https://www.ucd.ie/issda/data/).
 
@@
 # Early Results 
-```{r morefigs, out.width='90%', echo=FALSE, fig.cap='Permuted Null Distribution of Composite. Test statistic 28.89, P-value of P<0.001', fig.height=4.5}
+```{r morefigs, out.width='90%', echo=FALSE, fig.cap='Permuted Null Distribution of Composite. Test statistic 28.89, *p*-value of *p*<0.001', fig.height=4.5}
 continuous_permutation_function<-function(x){

### 2025-01-28 MSc_poster_template.Rmd
Citation added for EMERGE

@@
 When misused, they can reduce the power to detect treatment effect.
-This was seen in EMERGE, a randomised clinical trial investigating the effects of early initiation of Metformin treatment on developing gestational diabetes.
+This was seen in EMERGE [^1], a randomised clinical trial investigating the effects of early initiation of Metformin treatment on developing gestational diabetes.
 The composite was created using three measures of glycemic control; blood glucose levels of 5.1mmol/L or greater at gestation weeks 32 or 38, or initiation of Insulin treatment.
@@
 3.	To compare the permutation method with other approaches to analyzing composite outcomes
+
+[^1]: Dunne *et al,* 2023 doi:10.1001/jama.2023.19869
 
@@
 ggplot(squared_composite.df, aes(x= squared_composite))+
-  geom_histogram(bins = 200, fill = "salmon", color = "salmon")+
+  geom_histogram(bins = 200, fill = "mediumseagreen", color = "mediumseagreen")+
   scale_x_continuous(limits = c(0, 30)) +
@@
 
-* Compare the permutation test method with the approach Mc Menamin *et al.*, [^5] used to analyze composite outputs.
+* Compare the permutation test method with the approach Mc Menamin *et al.*, [^2] used to analyze composite outputs.
 * Compare different methods of creating the composite test statistic and their influence on power.
@@
 
-[^5]: Mc Menamin *et al,* 2018 doi:https://doi.org/10.1186/s13023-018-0819-1
+[^2]: Mc Menamin *et al,* 2018 doi:https://doi.org/10.1186/s13023-018-0819-1
 

### 2025-01-28 MSc_poster_template.Rmd
Minor adjustments

@@
 When misused, they can reduce the power to detect treatment effect.
-This was seen in EMERGE [^1], a randomised clinical trial investigating the effects of early initiation of Metformin treatment on developing gestational diabetes.
+This was seen in EMERGE, [^1] a randomised clinical trial investigating the effects of early initiation of Metformin treatment on developing gestational diabetes.
 The composite was created using three measures of glycemic control; blood glucose levels of 5.1mmol/L or greater at gestation weeks 32 or 38, or initiation of Insulin treatment.
@@
 
-This composite involves dichotomising blood glucose levels, greatly reducing power. Therefore, the objectives of this project are:
+This composite involves dichotomising blood glucose levels, greatly reducing power. Therefore, using data from EMERGE, the objectives of this project are:
 
-1.	To perform a meta-analysis of the primary outcome using permutations of the composite test statistic, and analyzing the glucose components as continuous variables
+1.	To perform a meta-analysis of the primary outcome using permutations of the composite test statistic, without dichotomising glucose variables
 

### 2025-01-29 MSc_poster_template.Rmd
Update MSc_poster_template.Rmd

@@
 
-```{r standard-plot, out.width='80%', fig.align='center', fig.cap='Great figure!', fig.height=4}
-ggplot(mpg, aes(displ, hwy, colour = class)) + 
-  geom_point()
+``````{r standard-plot, out.width='80%', fig.align='center', fig.cap='Great figure!', fig.height=4}
+# Load the necessary libraries
+library(ggplot2)
+library(hrbrthemes)
+library(readr)
+library(dplyr)
+library(reshape2)
+
+# Load the dataset
+data <- read.csv("~/Documents/data.csv")
+
+# Create a binary column for Insulin
+data$binary_insulin <- 0
+data$binary_insulin[data$InsulinYN == "Yes"] <- 1
+
+# Proportion test for Insulin usage between groups
+StudyInsulinTest <- prop.test(
+  x = c(
+    sum(data$InsulinYN[data$Group == "Metformin"] == "Yes", na.rm = TRUE),
+    sum(data$InsulinYN[data$Group == "Placebo"] == "Yes", na.rm = TRUE)
+  ),
+  n = c(
+    sum(data$Group == "Metformin"),
+    sum(data$Group == "Placebo")
+  )
+)
+
+# Perform t-tests for binary_insulin, gw32, and gw38
+insulin_t <- t.test(data$binary_insulin ~ data$Group)
+gw32_t <- t.test(data$gw32 ~ data$Group)
+gw38_t <- t.test(data$gw38 ~ data$Group)
+
+# Calculate composite scores
+y_squared_composite <- (insulin_t$statistic^2) + (gw32_t$statistic^2) + (gw38_t$statistic^2)
+study_non_squared_composite <- insulin_t$statistic + gw32_t$statistic + gw38_t$statistic
+
+# Visualization: Proportion of Insulin Usage by Group
+ggplot(data, aes(x = Group, fill = InsulinYN)) +
+  geom_bar(position = "fill") +
+  labs(
+    title = "Proportion of Insulin Usage by Group",
+    y = "Proportion",
+    x = "Group"
+  ) +
+  scale_fill_brewer(palette = "Set2") +
+  theme_minimal()
+
+# Visualization: Boxplots for gw32 and gw38 with cutoff line at 5.1 and p-values
+ggplot(data, aes(x = Group, y = gw32, fill = Group)) +
+  geom_boxplot() +
+  geom_hline(yintercept = 5.1, linetype = "dashed", color = "red") +
+  annotate("text", x = 1.5, y = max(data$gw32, na.rm = TRUE), 
+           label = paste("p-value:", round(gw32_t$p.value, 4)), 
+           color = "blue", size = 4) +
+  labs(
+    title = "Gestational Week 32 by Group",
+    y = "Week 32",
+    x = "Group",
+    caption = "Dashed red line represents the cutoff at 5.1"
+  ) +
+  theme_minimal()
+
+ggplot(data, aes(x = Group, y = gw38, fill = Group)) +
+  geom_boxplot() +
+  geom_hline(yintercept = 5.1, linetype = "dashed", color = "red") +
+  annotate("text", x = 1.5, y = max(data$gw38, na.rm = TRUE), 
+           label = paste("p-value:", round(gw38_t$p.value, 4)), 
+           color = "blue", size = 4) +
+  labs(
+    title = "Gestational Week 38 by Group",
+    y = "Week 38",
+    x = "Group",
+    caption = "Dashed red line represents the cutoff at 5.1"
+  ) +
+  theme_minimal()
+
+# Calculate and display correlation matrix
+correlation_vars <- data %>%
+  select(gw32, gw38, binary_insulin) 
+correlation_matrix <- cor(correlation_vars, use = "complete.obs", method = "pearson")
+
+# Plot correlation matrix as a heatmap
+correlation_melted <- melt(correlation_matrix)
+
+ggplot(correlation_melted, aes(Var1, Var2, fill = value)) +
+  geom_tile() +
+  geom_text(aes(label = round(value, 2)), color = "white") +
+  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0) +
+  labs(
+    title = "Correlation Matrix (Pearson)",
+    x = "Variables",
+    y = "Variables"
+  ) +
+  theme_minimal()
+
 ```

### 2025-01-29 MSc_poster_template.Rmd
Feed back for background addressed. Coding disparities for loading data addressed

--- 2025-01-28 MSc_poster_template.Rmd
@@
 # Background and Problem
-Composites are commonly used in clinical trials as the primary outcome.
-When misused, they can reduce the power to detect treatment effect.
-This was seen in EMERGE, [^1] a randomised clinical trial investigating the effects of early initiation of Metformin treatment on developing gestational diabetes.
-The composite was created using three measures of glycemic control; blood glucose levels of 5.1mmol/L or greater at gestation weeks 32 or 38, or initiation of Insulin treatment.
+Our project centres on  data from EMERGE, [^1] which was a randomised clinical trial investigating the effects of early initiation of Metformin treatment on developing gestational diabetes.
+The primary outcome of EMERGE was a composite created using three variables; blood glucose levels of 5.1mmol/L or greater at gestation weeks 32 or 38, or initiation of Insulin treatment.
 
@@
 
-This composite involves dichotomising blood glucose levels, greatly reducing power. Therefore, using data from EMERGE, the objectives of this project are:
+```{r}
 
-1.	To perform a meta-analysis of the primary outcome using permutations of the composite test statistic, without dichotomising glucose variables
+composite_metformin <- round(sum(tmpDat$compositeYN[tmpDat$Group == "Metformin"] == "Yes", na.rm = TRUE) /
+                               sum(!is.na(tmpDat$compositeYN[tmpDat$Group == "Metformin"])) * 100, 2)
+composite_placebo <- round(sum(tmpDat$compositeYN[tmpDat$Group == "Placebo"] == "Yes", na.rm = TRUE) /
+                             sum(!is.na(tmpDat$compositeYN[tmpDat$Group == "Placebo"])) * 100, 2)
+EMERGE_composite_p_value<-round(prop.test(x = c(sum(tmpDat$compositeYN[tmpDat$Group == "Metformin"] == "Yes", na.rm = T),
+                                                sum(tmpDat$compositeYN[tmpDat$Group== "Placebo"]== "Yes", na.rm = T)),
+                                          n = c(sum(!is.na(tmpDat$compositeYN[tmpDat$Group == "Metformin"])), sum(!is.na(tmpDat$compositeYN[tmpDat$Group == "Placebo"])))
+)$p.value, 2)
+EMERGE_background_stats<-c("Metformin" = composite_metformin, "Placebo" = composite_placebo,
+                           "P-value" = EMERGE_composite_p_value)
 
-2.	To compare various ways of creating the composite test statistic, and their power
+EMERGE_background_stats <- data.frame(t(as.matrix(EMERGE_background_stats)))
+knitr::kable(EMERGE_background_stats, caption = 'EMERGE *p*-values', align = 'c', format = "html") %>%
+  kable_styling(font_size = 20) %>%
+  add_header_above(c("Primary Composite Outcome N/total (%)" = 2, " " = 1))
+```
 
-3.	To compare the permutation method with other approaches to analyzing composite outcomes
+While the results of EMERGE were insignificant, the composite endpoint involved dichotomising blood glucose levels, greatly reducing power.
+Therefore, using data from EMERGE, the objectives of this project are:
+
+  1.	to investigate the potential of meta-analysis to improve power over standard analysis of composite endpoints when some of the components of the composite are continuous and some are binary
+
+2.	To compare various ways of creating the composite test statistic which is used in the meta-analysis, and their power
+
+3.	To compare the permutation-based meta analysis approach we have proposed here to other alternatives to standard analysis of composite endpoint
 
 [^1]: Dunne *et al,* 2023 doi:10.1001/jama.2023.19869
-
-# Data Sources and Datasets
-
-The syntax in this poster template and the `posterdown` package uses the same workflow approach as the R Markdown you know and love.
 
@@
 
-# Descriptive Statistics of Datasets
+```{r standard-plot, out.width='80%', fig.align='center', fig.cap='Great figure!', fig.height=4}
+# Load the necessary libraries
+library(ggplot2)
+library(hrbrthemes)
+library(readr)
+library(dplyr)
+library(reshape2)
 
-Usually you want to have a nice table displaying some important results that you have calculated.
-In `posterdown` this is as easy as using the `kable` table formatting you are probably use to as per typical R Markdown formatting.
+# Create a binary column for Insulin
+tmpDat$binary_insulin <- 0
+tmpDat$binary_insulin[tmpDat$InsulinYN == "Yes"] <- 1
 
-You can reference tables like so: Table \@ref(tab:mytable). Some basic summaries of the dataset are below:
+# Proportion test for Insulin usage between groups
+StudyInsulinTest <- prop.test(
+  x = c(
+    sum(tmpDat$InsulinYN[tmpDat$Group == "Metformin"] == "Yes", na.rm = TRUE),
+    sum(tmpDat$InsulinYN[tmpDat$Group == "Placebo"] == "Yes", na.rm = TRUE)
+  ),
+  n = c(
+    sum(tmpDat$Group == "Metformin"),
+    sum(tmpDat$Group == "Placebo")
+  )
+)
 
-```{r mytable, out.width='80%'}
-knitr::kable(iris[1:10, 1:4], caption = 'Table caption.',align = 'c',"html") %>%
-  kable_styling(font_size = 20)
-```
+# Perform t-tests for binary_insulin, gw32, and gw38
+insulin_t <- t.test(tmpDat$binary_insulin ~ tmpDat$Group)
+gw32_t <- t.test(tmpDat$gw32 ~ tmpDat$Group)
+gw38_t <- t.test(tmpDat$gw38 ~ tmpDat$Group)
 
-Figure \@ref(fig:standard-plot), and Figure \@ref(fig:morefigs) below show the patterns in our dataset.
-Make sure that all the details in your plots will be legible when printed (legend text, axis text, and any labels)
+# Calculate composite scores
+y_squared_composite <- (insulin_t$statistic^2) + (gw32_t$statistic^2) + (gw38_t$statistic^2)
+study_non_squared_composite <- insulin_t$statistic + gw32_t$statistic + gw38_t$statistic
 
-```{r standard-plot, out.width='80%', fig.align='center', fig.cap='Great figure!', fig.height=4}
-ggplot(mpg, aes(displ, hwy, colour = class)) + 
-  geom_point()
+# Visualization: Proportion of Insulin Usage by Group
+ggplot(tmpDat, aes(x = Group, fill = InsulinYN)) +
+  geom_bar(position = "fill") +
+  labs(
+    title = "Proportion of Insulin Usage by Group",
+    y = "Proportion",
+    x = "Group"
+  ) +
+  scale_fill_brewer(palette = "Set2") +
+  theme_minimal()
+
+# Visualization: Boxplots for gw32 and gw38 with cutoff line at 5.1 and p-values
+ggplot(tmpDat, aes(x = Group, y = gw32, fill = Group)) +
+  geom_boxplot() +
+  geom_hline(yintercept = 5.1, linetype = "dashed", color = "red") +
+  annotate("text", x = 1.5, y = max(tmpDat$gw32, na.rm = TRUE), 
+           label = paste("p-value:", round(gw32_t$p.value, 4)), 
+           color = "blue", size = 4) +
+  labs(
+    title = "Gestational Week 32 by Group",
+    y = "Week 32",
+    x = "Group",
+    caption = "Dashed red line represents the cutoff at 5.1"
+  ) +
+  theme_minimal()
+
+ggplot(tmpDat, aes(x = Group, y = gw38, fill = Group)) +
+  geom_boxplot() +
+  geom_hline(yintercept = 5.1, linetype = "dashed", color = "red") +
+  annotate("text", x = 1.5, y = max(tmpDat$gw38, na.rm = TRUE), 
+           label = paste("p-value:", round(gw38_t$p.value, 4)), 
+           color = "blue", size = 4) +
+  labs(
+    title = "Gestational Week 38 by Group",
+    y = "Week 38",
+    x = "Group",
+    caption = "Dashed red line represents the cutoff at 5.1"
+  ) +
+  theme_minimal()
+
+# Calculate and display correlation matrix
+correlation_vars <- tmpDat %>%
+  select(gw32, gw38, binary_insulin) 
+correlation_matrix <- cor(correlation_vars, use = "complete.obs", method = "pearson")
+
+# Plot correlation matrix as a heatmap
+correlation_melted <- melt(correlation_matrix)
+
+ggplot(correlation_melted, aes(Var1, Var2, fill = value)) +
+  geom_tile() +
+  geom_text(aes(label = round(value, 2)), color = "white") +
+  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0) +
+  labs(
+    title = "Correlation Matrix (Pearson)",
+    x = "Variables",
+    y = "Variables"
+  ) +
+  theme_minimal()
+
 ```

### 2025-01-29 MSc_poster_template.Rmd
Update MSc_poster_template.Rmd

@@
 ```
+# Analysis of EMERGE DATA SET
+The primary composite outcome at 32 and 38 weeks was not significantly different between the **metformin group (154 of 264) and the placebo group (82 of 262)**.
+This lack of significance is indicated by a **risk difference of -6.9%** (95% CI, -15.1% to 1.4%) and a**relative risk of 0.89** (95% CI, 0.78 to 1.02).
 

### 2025-01-29 MSc_poster_template.Rmd
Update MSc_poster_template.Rmd

@@
 poster_width: "841mm"
-column_numbers: 4
+column_numbers: 5
 column_margins: "0.5in"	

### 2025-01-29 MSc_poster_template.Rmd
Update MSc_poster_template.Rmd
idek
Update MSc_poster_template.Rmd
rit

@@
 library(ggplot2)
-load("\\Users\\brend\\Downloads\\data")
+tmpDat <- read.csv("/Users/athiramohan/Documents/data.csv")
 tmpDat$binary_insulin<-0
@@
 <!-- You can change where you put these breaks, and the "column_numbers: 4" argument in the YAMLS header to change the layout of the poster, with care! -->
-
-####
 

### 2025-01-29 MSc_poster_template.Rmd
Update MSc_poster_template.Rmd
fig =great remove

@@
 
-```{r standard-plot, out.width='80%', fig.align='center', fig.cap='Great figure!', fig.height=4}
+```{r standard-plot, out.width='80%', fig.align='center', fig.height=4}
 # Load the necessary libraries

### 2025-06-13 Code Repository.Rmd
Creating Neat collection of code

---
title: "Code Repository"
author: "Brendan Woods"
date: "`r Sys.Date()`"
output: html_document
---

### 2025-06-13 Code Repository.Rmd
Majority of code added

@@
 ---
+
+## Contents
+- Pre-processing
+- Creating the Composites using original EMERGE data
+- Permutation Code
+- Theoretical Distributions
+- Latent Variable Analysis McMenamin et al.
+- Power Estimation
+- Type I error rate Estimation
+- Simulating toy data set
+
+## Pre-processing
+```{r}
+library(dplyr)
+library(MASS)
+library(tidyr)
+library(mvtnorm)
+library(ggplot2)
+library(kableExtra)
+
+load("\\Users\\brend\\Downloads\\data")
+tmpDat <- tmpDat[!(is.na(tmpDat$InsulinYN) & is.na(tmpDat$gw32) & is.na(tmpDat$gw38)), ] ## removes individuals with no observations for 3 outcomes
+tmpDat$binary_insulin<-0
+tmpDat$binary_insulin[tmpDat$InsulinYN == "Yes"]<-1
+tmpDat$binary_group<-0
+tmpDat$binary_group[tmpDat$Group == "Metformin"]<-1
+tmpDat$binary_insulin<-as.numeric(tmpDat$binary_insulin)
+tmpDat$binary_EMERGE<-0
+tmpDat$binary_EMERGE[tmpDat$compositeYN == "Yes"]<-1
+```
+
+## Creating the Composites using original EMERGE data
+```{r}
+insulin_t_value<-t.test(tmpDat$binary_insulin[tmpDat$Group == "Metformin"], tmpDat$binary_insulin[tmpDat$Group == "Placebo"])$statistic
+gw32_t_value<-t.test(tmpDat$gw32[tmpDat$Group == "Metformin"], tmpDat$gw32[tmpDat$Group == "Placebo"])$statistic
+gw38_t_value<-t.test(tmpDat$gw38[tmpDat$Group == "Metformin"], tmpDat$gw38[tmpDat$Group == "Placebo"])$statistic
+
+study_squared_composite<-(insulin_t_value^2)+(gw32_t_value^2)+(gw38_t_value^2)
+study_non_squared_composite<-insulin_t_value+gw32_t_value+gw38_t_value
+
+```
+
+## Permutation Code
+Permutation Function
+```{r}
+permutation_function<-function(x, y, z){
+
+  permutation_summed_composite<-numeric(100000)
+  permutation_squared_composite <- numeric(100000)
+  i <-1
+
+  repeat{
+
+    tmpDat$Group<-sample(tmpDat$Group, nrow(tmpDat), replace = F)
+
+    t1 <- t.test(x[tmpDat$Group == "Metformin"], x[tmpDat$Group == "Placebo"])$statistic
+    t2 <- t.test(y[tmpDat$Group == "Metformin"], y[tmpDat$Group == "Placebo"])$statistic 
+    t3 <- t.test(z[tmpDat$Group == "Metformin"], z[tmpDat$Group == "Placebo"])$statistic
+
+    permutation_summed_composite[i] <- t1 + t2 +t3
+    permutation_squared_composite[i]<-  t1^2 + t2^2 + t3^2
+
+    i<- i +1
+
+    if(i > 100000) break}
+
+  list(permutation_summed_composite  = permutation_summed_composite,
+    permutation_squared_composite = permutation_squared_composite)
+
+}
+
+```
+
+Creating permuted null distributions using permutation function
+```{r}
+set.seed(24253739)
+permuted_null_distributions <-permutation_function(tmpDat$gw32, tmpDat$gw38, tmpDat$binary_insulin)
+permuted_null_distributions.df <-data.frame(permuted_null_distributions)
+```
+
+## Theoretical Distributions
+Summed test statistics  with equal weighting
+```{r}
+set.seed(24253739)
+multivariate <- thesis.df %>% dplyr::select(gw32, gw38, binary_insulin)
+sigma <- cor(multivariate, use = "complete.obs")
+w <- c(1, 1, 1)
+variance <-w%*%sigma%*%w
+size <- 1000000
+null_normal_distribution <- rnorm(size, mean = 0, sd = sqrt(variance))
+p_value <- sum(null_normal_distribution <= study_non_squared_composite)/size
+print(paste("P-value: ", p_value))
+
+null_normal_distribution.df <- data.frame(null_normal_distribution)
+
+```
+
+Plotting theoretical null distribution of summed test statistic composite
+```{r}
+ggplot(data = null_normal_distribution.df, aes(x =null_normal_distribution )) + geom_histogram(bins =250, colour = "skyblue", fill = "skyblue") + labs(title = "Normal Null Distribution", x = "", y = "Frequency") +
+  theme_minimal() +   geom_vline(xintercept = quantile(null_normal_distribution, probs = c(0.025, 0.975)), color = "black", linetype = "dashed", linewidth = 0.75) + 
+  geom_vline( xintercept = study_non_squared_composite, color = "gold", size = 1 ) + 
+  annotate("text", x = study_non_squared_composite, y = 10000, label = "EMERGE Composite", color = "gold", angle = 90, vjust = -0.5)
+```
+
+Squared test statistics with eigenvalue weighting
+```{r}
+set.seed(24253739)
+multivariate <- thesis.df %>% dplyr::select(gw32, gw38, binary_insulin)
+multivariate_cm <- cor(multivariate, use = "complete.obs")
+eigen_values <- eigen(multivariate_cm)$values
+size <- 1000000
+simulated_chi_sq_matrix <- matrix(rchisq(size * 3, df = 1), ncol = 3)
+eigen_values_chi_sq<- simulated_chi_sq_matrix %*%eigen_values  
+theoretical_null_squared <-data.frame(eigen_values_chi_sq)
+
+p_value <-sum(theoretical_null_squared >= study_squared_composite)/size
+print(paste("P-value: ", p_value))
+
+theoretical_null_squared <- data.frame(chi_sq = eigen_values_chi_sq)
+write.csv(theoretical_null_squared, "Theoretical Squared null distribution.csv", row.names = FALSE )
+
+```
+Plotting theoretical null distribution of squared test statistic composite with eigenvalue weights
+```{r}
+ggplot(data = theoretical_null_squared, aes(x =chi_sq )) + geom_histogram(bins =250, colour = "lightsalmon", fill = "lightsalmon") + labs(title = "", x = "", y = "") +
+  theme_minimal() + 
+  geom_vline(xintercept = quantile(theoretical_null_squared$chi_sq, probs = c(0.95)), color = "red", linetype = "dashed", linewidth = 0.75) +
+  annotate("text", x = 8.2, y = 25000, label = "95th Percentile", color = "red", angle = 90, vjust = -0.5, size = 5) +
+  theme(panel.grid.minor = element_blank())+
+  geom_vline( xintercept = study_squared_composite, color = "steelblue", size = 1 ) + 
+  annotate("text", x = study_squared_composite, y = 25000, label = "Composite", color = "steelblue", angle = 90, vjust = -0.5, size = 5)
+
+```
+
+## Latent Variable Analysis McMenamin et al.
+```{r}
+
+```
+
+##Power Estimation
+Power estimation function
+```{r}
+power_function <- function(x, y, z, e, n) {
+  summed_TS_power_result <- numeric(200)  
+  squared_TS_power_result <- numeric(200)  
+  EMERGE_composite_power <- numeric(200)
+  i <- 1
+  repeat {
+    placebo_group <- tmpDat[tmpDat$Group == "Placebo", ]
+    placebo_group <- placebo_group[complete.cases(placebo_group[, c(x, y, z)]), ]
+    sampled_placebo <- placebo_group[sample(nrow(placebo_group), n/2, replace = TRUE), ]
+
+    metformin_group <- tmpDat[tmpDat$Group == "Metformin", ]
+    metformin_group <- metformin_group[complete.cases(metformin_group[, c(x, y, z)]), ]
+    sampled_metformin <- metformin_group[sample(nrow(metformin_group), n/2, replace = TRUE), ]
+
+    t_test_result1<- t.test(sampled_metformin[x], sampled_placebo[x])
+    t_test_result2<- t.test(sampled_metformin[y], sampled_placebo[y])
+    t_test_result3<- t.test(sampled_metformin[z], sampled_placebo[z])
+    t_test_result4<- t.test(sampled_metformin[e], sampled_placebo[e])
+
+    summed_TS_power_result[i] <- (t_test_result1$statistic + t_test_result2$statistic + t_test_result3$statistic)
+    squared_TS_power_result[i]<-(((t_test_result1$statistic)^2)+((t_test_result2$statistic)^2)+((t_test_result3$statistic)^2))
+    EMERGE_composite_power[i] <- t_test_result4$p.value
+    i<- i +1
+    if(i > 200) break}
+  return(as.numeric(c(sum(summed_TS_power_result<quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025)| summed_TS_power_result > quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975))/ 200,## permutation test summed
+         sum(squared_TS_power_result>quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95))/ 200, ##permutation test squared
+         sum(squared_TS_power_result > quantile(theoretical_null_squared, 0.95))/ 200, ##asymptotic chi-square
+         sum((EMERGE_composite_power< 0.05))/ 200, ## EMERGE
+         sum(summed_TS_power_result<quantile(null_normal_distribution, 0.025)| summed_TS_power_result > quantile(null_normal_distribution, 0.975))/ 200)))
+}
+```
+
+Applying power estimation function
+```{r}
+sample_vector <- seq(from = 30, to = 540, by = 10)
+unified_power_results <- sapply(sample_vector, function(n_value) {
+  power_function("gw32", "gw38", "binary_insulin", "binary_EMERGE", n_value)
+})
+unified_power_results_df <-unified_power_results
+unified_power_results_df <- as.data.frame(t(unified_power_results)) 
+colnames(unified_power_results_df) <- c("Permuted: Summed TS", "Permuted: Squared TS", "Asymptotic Chi-square: Eigenvalue weights TS", "EMERGE Composite", "Asymptotic Normal: Equal weights TS")
+unified_power_results_df$sample_size <- sample_vector
+```
+
+Plotting power estimation
+```{r}
+power_100_latvar<-0.99
+power_150_latvar <- 0.99
+
+LVF_power_df <- data.frame( power = c( 0.99, 0.99),
+                                   sample_size = c(100, 150),
+                                   label = "Latent Variable")
+power_long <- unified_power_results_df %>%
+  pivot_longer(
+    cols = c("Permuted: Summed TS", "Permuted: Squared TS", "Asymptotic Chi-square: Eigenvalue weights TS", "EMERGE Composite", "Asymptotic Normal: Equal weights TS"),
+    names_to = "composite_type",
+    values_to = "power"
+  )
+write.csv(power_long, "power_long.csv", row.names = FALSE)
+
+ggplot(power_long, aes(x = sample_size, y = power, color = composite_type)) +
+  geom_line(size = 1, alpha = 0.5) +
+  labs(
+    title = expression(bold("")),
+    x = "Sample Size",
+    y = "Power"
+  ) +
+  scale_x_continuous(limits = c(30, NA), expand = c(0, 0)) +
+  scale_y_continuous(limits = c(0, 1.1)) +
+  theme_minimal() +
+  theme(
+    axis.title = element_text(size = 16),
+    axis.text = element_text(size = 12),
+    plot.title = element_text(size = 24),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 9, face = "bold"),
+    legend.position = "bottom",
+    panel.grid.minor = element_blank()
+  ) +
+  scale_color_manual(
+    values = c(
+      "Permuted: Summed TS" = "mediumseagreen",
+      "Permuted: Squared TS" = "blue",
+      "Asymptotic Chi-square: Eigenvalue weights TS" = "orange2",
+      "EMERGE Composite" = "magenta",
+      "Asymptotic Normal: Equal weights TS" = "firebrick"
+    )
+  ) + 
+  guides(color = guide_legend(nrow = 3)) + 
+  geom_point(data = LVF_power_df, aes(x = sample_size, y = power),
+    shape = 21,
+    fill = "black",
+    color = "black",
+    size = 2,
+    stroke = 1.2,
+    inherit.aes = FALSE,
+    vjust = -1
+  ) + 
+  geom_text(data = LVF_power_df, aes(x = sample_size, y = power, label = label),
+    vjust = -1,   
+    size = 2,
+    fontface = "bold",
+    color = "darkred",
+    inherit.aes = FALSE
+  )
+```
+
+## Type I error rate Estimation
+```{r}
+set.seed(24253739)
+type_I_error_function <- function(x, y, z, e, n) {
+  summed_TS_power_result <- numeric(10000)  
+  squared_TS_power_result <- numeric(10000)  
+  EMERGE_composite_power <- numeric(10000)
+  i <- 1
+  repeat {
+    sampled_data <- tmpDat[sample(nrow(tmpDat), n, replace = TRUE), ]
+    sampled_data$binary_g <- rep(c(0, 1), each = n / 2) ##binary_g refers to treatment group
+    sampled_metformin <- sampled_data[sampled_data$binary_g == 1,]
+    sampled_placebo <- sampled_data[sampled_data$binary_g == 0,]
+
+    t_test_result1<- t.test(sampled_metformin[x], sampled_placebo[x])
+    t_test_result2<- t.test(sampled_metformin[y], sampled_placebo[y])
+    t_test_result3<- t.test(sampled_metformin[z], sampled_placebo[z])
+    t_test_result4<- t.test(sampled_metformin[e], sampled_placebo[e])
+
+    summed_TS_power_result[i] <- (t_test_result1$statistic + t_test_result2$statistic + t_test_result3$statistic)
+    squared_TS_power_result[i] <- (((t_test_result1$statistic)^2)+((t_test_result2$statistic)^2)+((t_test_result3$statistic)^2))
+    EMERGE_composite_power[i] <- t_test_result4$p.value
+    i<- i +1
+    if(i > 10000) break}
+  return(c(sum(summed_TS_power_result<quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025) | summed_TS_power_result > quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975))/ 10000, ##permuted summed composite
+         sum(squared_TS_power_result>quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95))/ 10000, ##permutetd squared composite
+         sum(squared_TS_power_result > quantile(theoretical_null_squared, 0.95))/ 10000, ##theoretical squared composite
+         sum((EMERGE_composite_power< 0.05))/ 10000, ##EMERGE composite
+         sum(summed_TS_power_result<quantile(null_normal_distribution, 0.025)| summed_TS_power_result > quantile(null_normal_distribution, 0.975))/ 10000))
+}
+```
+
+Applying power estimation code and plotting
+```{r}
+sample_vector <- seq(from = 30, to = 540, by = 10)
+type_1_error_rate <- sapply(sample_vector, function(n_value) {
+  type_I_error_function("gw32", "gw38", "binary_insulin", "binary_EMERGE", n_value)
+})
+type_1_error_rate_df <- as.data.frame(t(type_1_error_rate)) 
+colnames(type_1_error_rate_df) <- c("Permuted: Summed TS", "Permuted: Squared TS", "Asymptotic Chi-square: Eigenvalue weights TS", "EMERGE Composite", "Asymptotic Normal: Equal weights TS")
+type_1_error_rate_df$sample_size <- sample_vector
+
+LVF_type_1_error_rate_100 <- 0.93
+LVF_type_1_error_rate_200 <- 0.96
+LVF_type_1_error_df <- data.frame( error_rate = c( 0.93, 0.96),
+                                   sample_size = c(100, 200),
+                                   label = "Latent Variable")
+
+error_long <- type_1_error_rate_df %>%
+  pivot_longer(
+    cols = c("Permuted: Summed TS", "Permuted: Squared TS", "Asymptotic Chi-square: Eigenvalue weights TS", "EMERGE Composite", "Asymptotic Normal: Equal weights TS"),
+    names_to = "composite_type",
+    values_to = "error_rate"
+  )
+
+ggplot(error_long, aes(x = sample_size, y = error_rate, color = composite_type)) +
+  geom_line(size = 1, alpha = 0.5) +
+  labs(
+    title = expression(bold("")),
+    x = "Sample Size",
+    y = "Type 1 error"
+  ) +
+  scale_x_continuous(limits = c(30, NA), expand = c(0, 0)) +
+  scale_y_continuous(limits = c(0, 0.1)) +
+  theme_minimal() +
+  theme(
+    axis.title = element_text(size = 12),
+    axis.text = element_text(size = 10),
+    plot.title = element_text(size = 28),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 11, face = "bold"),
+    legend.position = "bottom",
+    panel.grid.minor = element_blank()
+  ) +
+scale_color_manual(
+    values = c(
+      "Permuted: Summed TS" = "mediumseagreen",
+      "Permuted: Squared TS" = "blue",
+      "Asymptotic Chi-square: Eigenvalue weights TS" = "orange2",
+      "EMERGE Composite" = "magenta",
+      "Asymptotic Normal: Equal weights TS" = "firebrick"
+    )
+  ) + 
+  guides(color = guide_legend(nrow = 3)) +
+  geom_point(data = LVF_type_1_error_df, aes(x = sample_size, y = error_rate),
+    shape = 21,
+    fill = "black",
+    color = "black",
+    size = 2,
+    stroke = 1.2,
+    inherit.aes = FALSE,
+    vjust = -1
+  ) + 
+  geom_text(data = LVF_type_1_error_df, aes(x = sample_size, y = error_rate, label = label),
+    vjust = -1,   
+    size = 3.5,
+    fontface = "bold",
+    color = "darkred",
+    inherit.aes = FALSE
+  )
+
+```
+
+## Simulating toy data set
+
+pre-processing of tmpDat and acquiring parameters required for simulating data
+```{r}
+met_InsulinY_data <- filter(tmpDat, binary_insulin == 1, Group == "Metformin") 
+pla_InsulinY_data <- filter(tmpDat, binary_insulin == 1, Group == "Placebo")
+
+met_InsulinN_data <- filter(tmpDat, binary_insulin == 0, Group == "Metformin")
+pla_InsulinN_data <- filter(tmpDat, binary_insulin == 0, Group == "Placebo")
+
+met_correlation_matrix_InsulinY <- cor(x=dplyr::select(met_InsulinY_data, gw32, gw38), use = "pairwise.complete.obs")
+pla_correlation_matrix_InsulinY <- cor(x=dplyr::select(pla_InsulinY_data, gw32, gw38), use = "pairwise.complete.obs")
+
+met_correlation_matrix_InsulinN <- cor(x=dplyr::select(met_InsulinN_data, gw32, gw38), use = "pairwise.complete.obs")
+pla_correlation_matrix_InsulinN <- cor(x=dplyr::select(pla_InsulinN_data, gw32, gw38), use = "pairwise.complete.obs")
+
+avg_Met_gw32_InsulinY<-mean(tmpDat$gw32[tmpDat$Group == "Metformin" & tmpDat$binary_insulin == 1], na.rm=T) ##Average gw32 for metformin group with Insulin
+avg_Met_gw38_InsulinY<-mean(tmpDat$gw38[tmpDat$Group == "Metformin" & tmpDat$binary_insulin == 1], na.rm=T)## Same again for gw38
+
+avg_Met_gw32_InsulinN<-mean(tmpDat$gw32[tmpDat$Group == "Metformin" & tmpDat$binary_insulin == 0], na.rm=T)##Average gw38 for metformin with no Insulin
+avg_Met_gw38_InsulinN<-mean(tmpDat$gw38[tmpDat$Group == "Metformin" & tmpDat$binary_insulin == 0], na.rm=T)
+
+##Now for Placebo
+avg_Plcb_gw32_InsulinY<-mean(tmpDat$gw32[tmpDat$Group == "Placebo" & tmpDat$binary_insulin == 1], na.rm=T) 
+avg_Plcb_gw38_InsulinY<-mean(tmpDat$gw38[tmpDat$Group == "Placebo" & tmpDat$binary_insulin == 1], na.rm=T)
+
+avg_Plcb_gw32_InsulinN<-mean(tmpDat$gw32[tmpDat$Group == "Placebo" & tmpDat$binary_insulin == 0], na.rm=T)
+avg_Plcb_gw38_InsulinN<-mean(tmpDat$gw38[tmpDat$Group == "Placebo" & tmpDat$binary_insulin == 0], na.rm=T)
+```
+
+Generating the toy data set (aka dummy data)
+```{r}
+fake_met_insulinY <- mvrnorm(nrow(met_InsulinY_data), c( avg_Met_gw32_InsulinY, avg_Met_gw38_InsulinY), met_correlation_matrix_InsulinY)
+fake_met_insulinY <- data.frame(fake_met_insulinY)
+fake_met_insulinY$Group <- "Metformin"
+
+fake_pcb_insulinY <- mvrnorm(nrow(pla_InsulinY_data), c( avg_Plcb_gw32_InsulinY, avg_Plcb_gw38_InsulinY), pla_correlation_matrix_InsulinY)
+fake_pcb_insulinY <- data.frame(fake_pcb_insulinY)
+fake_pcb_insulinY$Group <- "Placebo"
+
+fake_met_insulinN <- mvrnorm(nrow(met_InsulinN_data), c( avg_Met_gw32_InsulinN, avg_Met_gw38_InsulinN), met_correlation_matrix_InsulinN)
+fake_met_insulinN <- data.frame(fake_met_insulinN)
+fake_met_insulinN$Group <- "Metformin"
+
+fake_pcb_insulinN <- mvrnorm(nrow(pla_InsulinN_data), c( avg_Plcb_gw32_InsulinN, avg_Plcb_gw38_InsulinN), pla_correlation_matrix_InsulinN)
+fake_pcb_insulinN <- data.frame(fake_pcb_insulinN)
+fake_pcb_insulinN$Group <- "Placebo"
+
+fake_InsulinY_df <- data.frame(rbind(fake_met_insulinY, fake_pcb_insulinY))
+fake_InsulinY_df$InsulinYN <- "Yes"
+fake_InsulinY_df$ID<- c(1:(nrow(met_InsulinY_data) + nrow(pla_InsulinY_data)))
+fake_InsulinY_df <- fake_InsulinY_df[, c("ID", "Group", "InsulinYN", "gw32", "gw38")]
+
+fake_InsulinN_df <- data.frame(rbind(fake_met_insulinN, fake_pcb_insulinN))
+fake_InsulinN_df$InsulinYN <- "No"
+fake_InsulinN_df$ID<- c((nrow(met_InsulinY_data) + nrow(pla_InsulinY_data) + 1): nrow(tmpDat))
+fake_InsulinN_df <- fake_InsulinN_df[, c("ID", "Group", "InsulinYN", "gw32", "gw38")]
+
+dummy_data<- rbind(fake_InsulinY_df, fake_InsulinN_df)
+
+dummy_data$binary_insulin<-0
+dummy_data$binary_insulin[dummy_data$InsulinYN == "Yes"]<-1
+
+dummy_data$binary_group<-0
+dummy_data$binary_group[dummy_data$Group == "Metformin"]<-1
+
+write.csv(dummy_data, "Simulated toy dataset.csv", rownames = FALSE)
+```
+
+table of parameters for dummy data
+```{r}
+dummy_data_parameters <- data.frame(group = c("Metformin, InsulinY", "Metformin, InsulinN",
+                                                 "Placebo, InsulinY", "Placebo, InsulinN"),
+                                       n = c(nrow(met_InsulinY_data), nrow(met_InsulinN_data),
+                                                  nrow(pla_InsulinY_data), nrow(pla_InsulinN_data)),
+                                       mean_gw32 = c(avg_Met_gw32_InsulinY, avg_Met_gw32_InsulinN,
+                                                     avg_Plcb_gw32_InsulinY, avg_Plcb_gw32_InsulinN),
+                                       mean_gw38 = c(avg_Met_gw38_InsulinY, avg_Met_gw38_InsulinN,
+                                                     avg_Plcb_gw38_InsulinY, avg_Plcb_gw38_InsulinN),
+                                       correlation = c(met_correlation_matrix_InsulinY[1,2], met_correlation_matrix_InsulinN[1,2],
+                                                       pla_correlation_matrix_InsulinY[1,2], pla_correlation_matrix_InsulinN[1,2]))
+colnames(dummy_data_parameters) <- c("Group", "N", "Mean GW32", "Mean GW38", "Correlation")
+dummy_data_parameters %>%
+  kable(
+    caption    = "Parameters of Simulated Toy Data",
+    booktabs   = TRUE,
+    digits     = c(0, 0, 2, 2, 2),
+    align      = c("l", "r", "r", "r", "r")
+  ) %>%
+  kable_styling(
+    latex_options = c("striped", "hold_position"), 
+    full_width    = FALSE
+  )
+```
+
+Toy data adapted for McMenamins app
+Require baseline gw32 and gw38 values, which we set as 0 for all except two individuals, one of which has 0.1 as the baseline value for gw32 or gw38
+Additionally, the values had to be "inverted" for the mcmenamin app to work properly. This was assessed by matching the risk difference calculated by the app to that reported in the EMERGE paper.
+```{r}
+simulated_data <- dummy_data %>% dplyr::select(patient_id = ID, treat = binary_group, Y1 = gw32, Y2 = gw38, Ybin = binary_insulin)
+simulated_data$Y1_0 <- 0
+simulated_data$Y2_0 <- 0
+simulated_data$Y1_0[1] <- -0.1
+simulated_data$Y2_0[2] <- -0.1
+inverted_data$Ybin <- ifelse(simulated_data$Ybin == 1, 0, 1)
+simulated_data$Y1 <- simulated_data$Y1 * -1
+simulated_data$Y2 <- simulated_data$Y2 * -1
+write.csv(simulated_data, "Simulated Inverted Emerge Data.csv", row.names = FALSE)
+
+```

### 2025-06-13 MSc_poster_template.Rmd
Trying to fix merge issues

--- 2025-01-29 MSc_poster_template.Rmd
@@
 poster_width: "841mm"
-column_numbers: 5
+column_numbers: 4
 column_margins: "0.5in"	
@@
 
-```{r}
+```{r mytable, out.width='80%'}
 
@@
 EMERGE_background_stats <- data.frame(t(as.matrix(EMERGE_background_stats)))
-knitr::kable(EMERGE_background_stats, caption = 'EMERGE *p*-values', align = 'c', format = "html") %>%
+styled_caption <- "<span style='font-family: Arial; font-size: 18px; font-weight: bold; color: black;'>Summary of Primary Composite in EMERGE</span>"
+knitr::kable(EMERGE_background_stats, caption = styled_caption, align = 'c', format = "html") %>%
   kable_styling(font_size = 20) %>%
@@
 ```
-# Analysis of EMERGE DATA SET
-The primary composite outcome at 32 and 38 weeks was not significantly different between the **metformin group (154 of 264) and the placebo group (82 of 262)**.
-This lack of significance is indicated by a **risk difference of -6.9%** (95% CI, -15.1% to 1.4%) and a**relative risk of 0.89** (95% CI, 0.78 to 1.02).
 
@@
 #### 
+# Descriptive Statistics of Datasets
 
@@
 
-# Create a binary column for Insulin
-tmpDat$binary_insulin <- 0
-tmpDat$binary_insulin[tmpDat$InsulinYN == "Yes"] <- 1
-
-# Proportion test for Insulin usage between groups
-StudyInsulinTest <- prop.test(
-  x = c(
-    sum(tmpDat$InsulinYN[tmpDat$Group == "Metformin"] == "Yes", na.rm = TRUE),
-    sum(tmpDat$InsulinYN[tmpDat$Group == "Placebo"] == "Yes", na.rm = TRUE)
-  ),
-  n = c(
-    sum(tmpDat$Group == "Metformin"),
-    sum(tmpDat$Group == "Placebo")
-  )
-)
-
 # Perform t-tests for binary_insulin, gw32, and gw38
@@
 gw38_t <- t.test(tmpDat$gw38 ~ tmpDat$Group)
-
-# Calculate composite scores
-y_squared_composite <- (insulin_t$statistic^2) + (gw32_t$statistic^2) + (gw38_t$statistic^2)
-study_non_squared_composite <- insulin_t$statistic + gw32_t$statistic + gw38_t$statistic
-
-# Visualization: Proportion of Insulin Usage by Group
-ggplot(tmpDat, aes(x = Group, fill = InsulinYN)) +
-  geom_bar(position = "fill") +
-  labs(
-    title = "Proportion of Insulin Usage by Group",
-    y = "Proportion",
-    x = "Group"
-  ) +
-  scale_fill_brewer(palette = "Set2") +
-  theme_minimal()
 
@@
     y = "Week 32",
-    x = "Group",
-    caption = "Dashed red line represents the cutoff at 5.1"
+    x = "Group"
   ) +
-  theme_minimal()
+  theme_minimal()+
+  theme(axis.title = element_text( size = 24),
+        axis.text = element_text(size = 18),
+        plot.title = element_text(size = 28))
 
@@
     y = "Week 38",
-    x = "Group",
-    caption = "Dashed red line represents the cutoff at 5.1"
+    x = "Group"
   ) +
-  theme_minimal()
+  theme_minimal()+
+  theme(axis.title = element_text( size = 24),
+        axis.text = element_text(size = 18),
+        plot.title = element_text(size = 28))
 
@@
   ) +
-  theme_minimal()
+  theme_minimal()+
+  theme(axis.title = element_text( size = 24),
+        axis.text = element_text(size = 18),
+        plot.title = element_text(size = 28))
 
@@
 
-```{r morefigs2, out.width='95%', echo=FALSE, fig.cap='Power of Permutation Test', fig.height=4.5}
+```{r morefigs2, out.width='95%', echo=FALSE, fig.cap='Power of Permutation Test', fig.height=4}
 

### 2025-06-14 Code Repository.Rmd
Some adjustments

@@
 study_squared_composite<-(insulin_t_value^2)+(gw32_t_value^2)+(gw38_t_value^2)
-study_non_squared_composite<-insulin_t_value+gw32_t_value+gw38_t_value
+study_summed_composite<-insulin_t_value+gw32_t_value+gw38_t_value
 
@@
 permuted_null_distributions.df <-data.frame(permuted_null_distributions)
+write.csv(permuted_null_distributions.df, "Permuted Null Distributions.csv", row.names = FALSE)
+
+```
+
+Plotting permuted null distributions
+Summed test statistics composite
+```{r warning=FALSE}
+ggplot(permuted_null_distributions.df, aes(x= permutation_summed_composite))+
+  geom_histogram(bins = 200, fill = "lightsalmon", color = "lightsalmon")+
+  scale_x_continuous(limits = c(-10, 10)) +
+  labs(x = "Test Statistic", y = "Frequency") +
+  geom_vline(xintercept = quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025), color = "red", linetype = "dashed", size = 0.75)+
+    geom_vline(xintercept = quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975), color = "red", linetype = "dashed", size = 0.75)+
+  annotate("text", x = quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025), y = 1250, label = "2.5th Percentile", color = "red", angle = 90, vjust = -0.5, size = 5)+
+    annotate("text", x = quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975), y = 1250, label = "97.5th Percentile", color = "red", angle = 90, vjust = -0.5, size = 5)+
+  theme_minimal()+ theme(panel.grid.minor = element_blank()) +
+  geom_vline( xintercept = study_summed_composite, color = "steelblue", size = 1 )+
+  annotate("text", x = study_summed_composite, y = 1250, label = "Composite", color = "steelblue", angle = 90, vjust = -0.5, size = 5) + theme(axis.title = element_text(size = 16))
+```
+
+Permuted Squared test statistics composite
+```{r warning=FALSE}
+ggplot(permuted_null_distributions.df, aes(x= permutation_squared_composite))+
+  geom_histogram(bins = 200, fill = "darksalmon", color = "darksalmon")+
+  scale_x_continuous(limits = c(0, 30)) +
+  labs(x = "Test Statistic", y = "Frequency") +
+  geom_vline(xintercept = quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95), color = "red", linetype = "dashed", size = 0.75)+
+  annotate("text", x = quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95), y = 2500, label = "95th Percentile", color = "red", angle = 90, vjust = -0.5)+
+  theme_minimal()+ theme(panel.grid.minor = element_blank(),
+                         axis.title = element_text(size = 16)) +
+  geom_vline(xintercept = study_squared_composite, color = "steelblue", size = 1 ) +
+  annotate("text", x = study_squared_composite, y = 2500, label = "Composite", color = "steelblue", angle = 90, vjust = -0.5)
 ```
@@
 set.seed(24253739)
-multivariate <- thesis.df %>% dplyr::select(gw32, gw38, binary_insulin)
+multivariate <- tmpDat %>% dplyr::select(gw32, gw38, binary_insulin)
 sigma <- cor(multivariate, use = "complete.obs")
@@
 null_normal_distribution <- rnorm(size, mean = 0, sd = sqrt(variance))
-p_value <- sum(null_normal_distribution <= study_non_squared_composite)/size
+p_value <- sum(null_normal_distribution <= study_summed_composite)/size
 print(paste("P-value: ", p_value))
@@
 null_normal_distribution.df <- data.frame(null_normal_distribution)
+write.csv(null_normal_distribution.df, "Theoretical Summed null distribution.csv", row.names = FALSE )
 
@@
   theme_minimal() +   geom_vline(xintercept = quantile(null_normal_distribution, probs = c(0.025, 0.975)), color = "black", linetype = "dashed", linewidth = 0.75) + 
-  geom_vline( xintercept = study_non_squared_composite, color = "gold", size = 1 ) + 
-  annotate("text", x = study_non_squared_composite, y = 10000, label = "EMERGE Composite", color = "gold", angle = 90, vjust = -0.5)
+  geom_vline( xintercept = study_summed_composite, color = "gold", size = 1 ) + 
+  annotate("text", x = study_summed_composite, y = 10000, label = "EMERGE Composite", color = "gold", angle = 90, vjust = -0.5)
 ```
@@
 set.seed(24253739)
-multivariate <- thesis.df %>% dplyr::select(gw32, gw38, binary_insulin)
+multivariate <- tmpDat %>% dplyr::select(gw32, gw38, binary_insulin)
 multivariate_cm <- cor(multivariate, use = "complete.obs")
@@
 
-##Power Estimation
+## Power Estimation
 Power estimation function
@@
 power_function <- function(x, y, z, e, n) {
-  summed_TS_power_result <- numeric(200)  
-  squared_TS_power_result <- numeric(200)  
-  EMERGE_composite_power <- numeric(200)
+  summed_TS_power_result <- numeric(1000)  
+  squared_TS_power_result <- numeric(1000)  
+  EMERGE_composite_power <- numeric(1000)
   i <- 1
@@
     i<- i +1
-    if(i > 200) break}
-  return(as.numeric(c(sum(summed_TS_power_result<quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025)| summed_TS_power_result > quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975))/ 200,## permutation test summed
-         sum(squared_TS_power_result>quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95))/ 200, ##permutation test squared
-         sum(squared_TS_power_result > quantile(theoretical_null_squared, 0.95))/ 200, ##asymptotic chi-square
-         sum((EMERGE_composite_power< 0.05))/ 200, ## EMERGE
-         sum(summed_TS_power_result<quantile(null_normal_distribution, 0.025)| summed_TS_power_result > quantile(null_normal_distribution, 0.975))/ 200)))
+    if(i > 1000) break}
+  return(as.numeric(c(sum(summed_TS_power_result<quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025)| summed_TS_power_result > quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975))/ 1000,## permutation test summed
+         sum(squared_TS_power_result>quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95))/ 1000, ##permutation test squared
+         sum(squared_TS_power_result > quantile(theoretical_null_squared$chi_sq, 0.95))/ 1000, ##asymptotic chi-square
+         sum((EMERGE_composite_power< 0.05))/ 1000, ## EMERGE
+         sum(summed_TS_power_result<quantile(null_normal_distribution, 0.025)| summed_TS_power_result > quantile(null_normal_distribution, 0.975))/ 1000)))
 }
@@
 unified_power_results_df$sample_size <- sample_vector
+
 ```
@@
          sum(squared_TS_power_result>quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95))/ 10000, ##permutetd squared composite
-         sum(squared_TS_power_result > quantile(theoretical_null_squared, 0.95))/ 10000, ##theoretical squared composite
+         sum(squared_TS_power_result > quantile(theoretical_null_squared$chi_sq, 0.95))/ 10000, ##theoretical squared composite
          sum((EMERGE_composite_power< 0.05))/ 10000, ##EMERGE composite
@@
   )
+write.csv(error_long, "Type I error estimation.csv", row.names = FALSE)
 
@@
 ```{r}
+tmpDat <- tmpDat[(!is.na(tmpDat$InsulinYN) & !is.na(tmpDat$gw32) & !is.na(tmpDat$gw38)), ]
 met_InsulinY_data <- filter(tmpDat, binary_insulin == 1, Group == "Metformin") 
@@
 
-write.csv(dummy_data, "Simulated toy dataset.csv", rownames = FALSE)
+write.csv(dummy_data, "Simulated toy dataset.csv", row.names = FALSE)
 ```
@@
 simulated_data$Y2_0[2] <- -0.1
-inverted_data$Ybin <- ifelse(simulated_data$Ybin == 1, 0, 1)
+simulated_data$Ybin <- ifelse(simulated_data$Ybin == 1, 0, 1)
 simulated_data$Y1 <- simulated_data$Y1 * -1

### 2025-06-15 Code Repository.Rmd
Updated Latent variable code
Latent variable code added

@@
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(
+  echo    = FALSE,
+  message = FALSE,
+  warning = FALSE)
+
+```
 
@@
 library(kableExtra)
+library(stats)
+library(nlme)
+library(boot)
+library(matrixcalc)
+library(numDeriv)
+library(cubature)
+library(optimx)
+library(brglm)
+library(Matrix)
+library(matrixcalc)
 
@@
 write.csv(permuted_null_distributions.df, "Permuted Null Distributions.csv", row.names = FALSE)
-
+p_value_squared_composite <- sum(permuted_null_distributions.df$permutation_squared_composite >= study_squared_composite)/100000
+p_value_summed_composite <- sum(permuted_null_distributions.df$permutation_summed_composite <= study_summed_composite)/100000
+print(paste("P-value squared composite: ", p_value_squared_composite))
+print(paste("P-value summed composite: ", p_value_summed_composite))
 ```
@@
 ## Latent Variable Analysis McMenamin et al.
+The following code is from McMenamin et al. 2021 which has been modified to remove the code for baseline variables of the continuous outcomes.
 ```{r}
+#Likelihood function
 
+f<-function(X,dat)
+{
+  #data
+  #dat<-dat[!is.na(dat[,3]),]#change this to baseline measure when it is added?
+
+  #parameters
+  alpha0 <- X[1]
+    alpha1 <- X[2]
+       beta0 <- X[3]
+         beta1 <- X[4]
+           psi0 <- X[5]
+             psi1 <- X[6]
+
+  #covariance parameters
+  sig1 <- exp(X[7]) 
+    sig2 <- exp(X[8]) 
+     rho12 <- 2*inv.logit(X[9])-1
+       rho13 <- 2*inv.logit(X[10])-1
+          rho23 <- 2*inv.logit(X[11])-1
+
+  #Known cutoffs
+  tau0 <- -Inf
+   tau1 <- 0
+     tau2 <- +Inf
+
+  print(X)
+
+  #model means
+  muz1<-alpha0+alpha1*dat[,2]
+   muz2<-beta0+beta1*dat[,2]
+     muz3<-psi0+psi1*dat[,2]
+
+  #conditional means
+  muz3cond <- muz3+((rho13-rho12*rho23)*(dat[,3]-muz1)/(sqrt(sig1)*(1-(rho12)^2)))+((rho23-rho13*rho12)*(dat[,4]-muz2)/(sqrt(sig2)*(1-(rho12)^2)))
+
+  #conditional covariance
+  sigcond <- 1-(((rho13)^2-2*rho12*rho13*rho23+(rho23)^2)/(1-(rho12)^2))
+
+  #continuous bivariate covariance
+  matbiv11 <- (sig1)^2
+   matbiv12 <- rho12*sig1*sig2
+    matbiv22 <- (sig2)^2
+      Sigbiv   <- matrix(c(matbiv11, matbiv12, matbiv12, matbiv22), nrow=2, ncol=2)#check
+        Sigbiv <- (Sigbiv*t(Sigbiv))^0.5
+
+  #upperlimits
+  mulim0<-matrix( tau1-muz3cond, ncol=1)
+   mulim1<-matrix( tau2-muz3cond, ncol=1)
+
+  #binprob
+  pr30<-apply(mulim0,1,function(x){return(pmvnorm(lower=-Inf,upper=x,mean=0,sigma = sigcond))})
+   pr31<-apply(mulim1,1,function(x){return(pmvnorm(lower=-Inf,upper=x,mean=0,sigma = sigcond))})
+     prz12<-dmvnorm(cbind(dat[,3],dat[,4]), c(mean(muz1), mean(muz2)), Sigbiv)
+
+  #Likelihood
+
+  #components of likelihood,k=0,1 (binary)
+  l0<-log(pr30)+log(prz12)#k=0
+  l1<-log(pr31-pr30)+log(prz12)#k=1
+
+  data0 <- cbind(dat[,5],l0)#0
+   data1 <- cbind(dat[,5],l1)#1
+
+  #0
+  data0[data0[,1]==1,2]<-0
+
+  #1
+  data1[data1[,1]==0,2]<-0
+
+  t0 <- sum(data0[,2])
+    t1 <- sum(data1[,2])
+
+  #-log(likelihood)
+  Tfinal<-sum(t0)+sum(t1)
+
+  return(-Tfinal)
+}
+
+lowerlim <- c(-Inf,-Inf,-Inf,-Inf,-Inf,-Inf,-Inf,-Inf,-Inf,-Inf,-Inf)
+upperlim <- c(+Inf,+Inf,+Inf,+Inf,+Inf,+Inf,+Inf,+Inf,+Inf,+Inf,+Inf)
+
+##Probability of response
+integrand<-function(Zint,meantreat,meanuntreat,mle)
+{
+
+  sigmahat=matrix(nrow=3,ncol=3)
+    sigmahat[1,1]=(exp(mle[7]))^2
+     sigmahat[2,1]=(2*inv.logit(mle[9])-1)*(exp(mle[7]))*exp(mle[8])
+       sigmahat[3,1]=(2*inv.logit(mle[10])-1)*(exp(mle[7]))
+           sigmahat[1,2]=sigmahat[2,1]
+             sigmahat[2,2]=(exp(mle[8]))^2
+               sigmahat[3,2]=(2*inv.logit(mle[11])-1)*(exp(mle[8]))
+                 sigmahat[1,3]=sigmahat[3,1]
+                   sigmahat[2,3]=sigmahat[3,2]
+                     sigmahat[3,3]=1
+
+  xtreat<-cbind(-meantreat[,1]+Zint[1], -meantreat[,2]+Zint[2], -meantreat[,3]+Zint[3])
+    xuntreat<-cbind(-meanuntreat[,1]+Zint[1],-meanuntreat[,2]+Zint[2],-meanuntreat[,3]+Zint[3])
+
+  pdftreat=dmvnorm(xtreat, mean=c(0,0,0),sigma=sigmahat)
+     pdfuntreat=dmvnorm(xuntreat, mean=c(0,0,0),sigma=sigmahat)
+
+  return(c(mean(pdftreat),mean(pdfuntreat)))
+}
+
+probofsuccess<-function(mle,n,dat,eta)
+{
+  n=n
+
+  meantreat=cbind(cbind(rep(1,n),rep(1,n))%*%c(mle[1:2]),cbind(rep(1,n),rep(1,n))%*%c(mle[3:4]), 
+                  cbind(rep(1,n),rep(1,n))%*%mle[5:6])   
+  meanuntreat=cbind(cbind(rep(1,n),rep(0,n))%*%c(mle[1:2]),cbind(rep(1,n),rep(0,n))%*%c(mle[3:4]), 
+                    cbind(rep(1,n),rep(0,n))%*%mle[5:6])
+
+  #lower and upper bounds
+  minmean1=min(c(meantreat[,1],meanuntreat[,1]))
+  minmean2=min(c(meantreat[,2],meanuntreat[,2]))
+  minmean3=min(c(meantreat[,3],meanuntreat[,3]))
+
+  maxmean1=max(c(meantreat[,1],meanuntreat[,1]))
+  maxmean2=max(c(meantreat[,2],meanuntreat[,2]))
+  maxmean3=max(c(meantreat[,3],meanuntreat[,3]))
+
+  lower=c(qnorm(1e-15,minmean1,exp(mle[7])),qnorm(1e-15,minmean2,exp(mle[8])),qnorm(1e-15,minmean3,1))
+
+  upper=c(eta[1],eta[2],0)
+
+  a=cuhre(f=integrand,nComp=2,lower=lower,upper=upper,flags=list(verbose=0,final=1,pseudo.random=0,mersenne.seed=NULL),
+          meantreat=meantreat,meanuntreat=meanuntreat,mle=mle)
+  #return(c(a$value[1],a$value[2]))
+  #return(c(a$value[1]-a$value[2],a$value[1],a$value[2]))
+  return(c((log(a$integral[1]/(1-a$integral[1]))-log(a$integral[2]/(1-a$integral[2]))),log(a$integral[1]/a$integral[2]),
+           a$integral[1]-a$integral[2],a$integral[1],a$integral[2]))
+  #return(log(a$value[1]/a$value[2]))
+}
+
+#Partial derivatives
+
+partials<-function(mle,n,dat,eta)
+{
+  p=length(mle)
+  fit1<-probofsuccess(mle,n,dat,eta)
+  fitOR<-fit1[1]
+  fitRR<-fit1[2]
+  fitRD<-fit1[3]
+  partials.augbinOR<-as.vector(rep(0,p))
+  partials.augbinRR<-as.vector(rep(0,p))
+  partials.augbinRD<-as.vector(rep(0,p))
+
+  for(i in 1:p){
+    valueupdate=mle
+    valueupdate[i]=valueupdate[i]+0.000001
+
+    updateprobOR=probofsuccess(valueupdate,n,dat,eta)[1]
+    updateprobRR=probofsuccess(valueupdate,n,dat,eta)[2]
+    updateprobRD=probofsuccess(valueupdate,n,dat,eta)[3]
+
+    partials.augbinOR[i]=(updateprobOR-fitOR)/0.000001
+    partials.augbinRR[i]=(updateprobRR-fitRR)/0.000001
+    partials.augbinRD[i]=(updateprobRD-fitRD)/0.000001
+
+  }
+
+  return(c(partials.augbinOR,partials.augbinRR,partials.augbinRD,fit1))
+}
+
+result.latent<-as.list(NULL)
+results<-as.list(NULL)
+
+#Function for treatment effects and CIs for both methods
+
+LatVarfunc<-function(dat,eta){
+  n=dim(dat)[1]
+
+  #Starting values
+  lm1<-lm(dat[,3]~dat[,2])
+  lm2<-lm(dat[,4]~dat[,2])
+  lm3<-lm(dat[,5]~dat[,2])
+  sig1est<-log(sqrt(var(dat[,3])))
+  sig2est<-log(sqrt(var(dat[,4])))
+  rho12est<-log(((cor(dat[,3],dat[,4])+1)/2)/(1-(cor(dat[,3],dat[,4])+1)/2))
+  rho13est<-log(((cor(dat[,3],dat[,5])+1)/2)/(1-(cor(dat[,3],dat[,5])+1)/2))
+  rho23est<-log(((cor(dat[,4],dat[,5])+1)/2)/(1-(cor(dat[,4],dat[,5])+1)/2))
+
+  X<-c(lm1$coef[1],lm1$coef[2],lm2$coef[1],lm2$coef[2],lm3$coef[1],lm3$coef[2],sig1est,sig2est,rho12est, rho13est,rho23est)
+  X<-as.vector(X)
+
+  #Latent variable
+
+  mlefit=optimx(X,f,dat=dat,lower=lowerlim,upper=upperlim,method="nlminb",control=list(rel.tol=1e-12))
+  mle<-coef(mlefit[1,])
+  hess<-attr(mlefit,"details")["nlminb",]$nhatend
+  print(hess)
+  mlecov=ginv(hess)
+  mlecov<-nearPD(mlecov)$mat
+  se<-sqrt(diag(mlecov))
+  print(se)
+
+  part<-partials(mle,n,dat,eta)
+  print(part)
+  meanOR<-part[34]
+
+  partsOR <- part[1:11]
+  varianceOR=t(partsOR)%*%mlecov%*%partsOR
+  varianceOR=varianceOR[1,1]
+
+  meanRR<-part[35]
+  partsRR<-part[12:22]
+  varianceRR=t(partsRR)%*%mlecov%*%partsRR
+  varianceRR=varianceRR[1,1]
+
+  meanRD<-part[36]
+  partsRD<-part[23:33]
+  varianceRD=t(partsRD)%*%mlecov%*%partsRD
+  varianceRD=varianceRD[1,1]
+
+  CIOR<-c(meanOR-1.96*sqrt(varianceOR),meanOR,meanOR+1.96*sqrt(varianceOR))
+  CIRR<-c(meanRR-1.96*sqrt(varianceRR),meanRR,meanRR+1.96*sqrt(varianceRR))
+  CIRD<-c(meanRD-1.96*sqrt(varianceRD),meanRD,meanRD+1.96*sqrt(varianceRD))
+
+  probresplat<-c(part[37],part[38])
+  result.latent<-c(CIOR,CIRR,CIRD,probresplat)
+
+  epshat1 <- dat[,3]-(mle[1]+mle[2]*dat[,2])
+  epshat2 <- dat[,4]-(mle[3]+mle[4]*dat[,2])
+  epshat3 <- dat[,5]-(mle[5]+mle[6]*dat[,2])
+  epshat <- as.matrix(cbind(epshat1,epshat2,epshat3))
+
+  sig1hat <- exp(mle[7]) 
+  sig2hat <- exp(mle[8]) 
+  ##Correlation coefficients (rho12, rho13, rho23) must lie in the interval (-1, 1).
+However, optimization routines like optimx work best on    unconstrained parameters, ranging from -Inf to +Inf.
+To enable this, the code uses the logistic (sigmoid) transformation to map an           unconstrained value to the range (0,1), and then linearly transforms it to (−1,1).##
+  rho12hat <- 2*inv.logit(mle[9])-1
+  rho13hat <- 2*inv.logit(mle[10])-1
+  rho23hat <- 2*inv.logit(mle[11])-1
+
+  SigHat<-matrix(data=c(sig1hat^2,rho12hat*sig1hat*sig2hat,rho13hat*sig1hat,rho12hat*sig1hat*sig2hat,sig2hat^2,
+           rho23hat*sig2hat,rho13hat*sig1hat,rho23hat*sig2hat,1),ncol=3)
+  modres <- diag((epshat)%*%ginv(SigHat)%*%t(epshat))
+
+  results<-c(result.latent,modres,mle,se)
+  results1 <- c(result.latent, mle, se)
+  return(results1)
+}
+```
+
+The output of the latvarfunc is as follows:
+1	logOR_LCI	Lower 95% CI for the log odds ratio
+2	logOR	Point estimate for the log odds ratio
+3	logOR_UCI	Upper 95% CI for the log odds ratio
+4	logRR_LCI	Lower 95% CI for the log relative risk
+5	logRR	Point estimate for the log relative risk
+6	logRR_UCI	Upper 95% CI for the log relative risk
+7	RD_LCI	Lower 95% CI for the risk difference
+8	RD	Point estimate for the risk difference
+9	RD_UCI	Upper 95% CI for the risk difference
+
+Reprocessing tmpDat into the suitable format "inverted_data" for the LatVarfunc
+Requires setting columns in a specific order and inverting the data
+Column 1 is patient ID
+Column 2 is treatment (binary 0,1)
+Column 3 is gw32
+Column 4 is gw38
+Column 5 is Insulin (binary 0, 1)
+Make sure to remove ## for the code to run
+```{r}
+data <- tmpDat[,c("Label", "binary_group", "gw32", "gw38", "binary_insulin")]
+inverted_data <- data[complete.cases(data), ]
+inverted_data$binary_insulin <- ifelse(inverted_data$binary_insulin == 1, 0, 1)
+inverted_data$gw32 <- inverted_data$gw32 * -1
+inverted_data$gw38 <- inverted_data$gw38 * -1
+inverted_eta <- c(-5.1, -5.1) ## these are the "cut-off" points for the continuous variables
+results <- LatVarfunc(inverted_data, inverted_eta)
+names(results) <- c("logOR_LCI", "logOR", "logOR_UCI")
+```
+
+Power estimation of Latent variable analysis
+*Note that this takes a long time to run, and as such we only ran it for two sample sizes, 100 and 150*
+```{r}
+set.seed(24253739)
+power_function_null_dist <- function(n) {
+  power_result <- matrix(NA, nrow = 100, ncol = 3) 
+  i <- 1
+  repeat {
+    placebo_group <- inverted_data[inverted_data$binary_group == 0, ]
+    sampled_placebo <- placebo_group[sample(nrow(placebo_group), n/2, replace = TRUE), ]
+
+    metformin_group <- inverted_data[inverted_data$binary_group == 1, ]
+    sampled_metformin <- metformin_group[sample(nrow(metformin_group), n/2, replace = TRUE), ]
+
+    sampled_data <- rbind(sampled_placebo, sampled_metformin) ##Needed for latvarfunc?
+    inverted_eta <- c(-5.1, -5.1)
+
+    LatVarfunc_result1<- LatVarfunc(sampled_data, inverted_eta)
+
+    power_result[i, ] <- LatVarfunc_result1[1:3]
+    i<- i +1
+    if(i > 100) break}
+  return(sum(power_result[,1] > 0 | power_result[,3] < 0)/ 100)
+}
+sample_vector <- 100
+##null_dist_summed_TS_results <- sapply(sample_vector, function(n_value) {power_function_null_dist(n_value)})
+power_100_latvar<-0.99
+power_150_latvar <- 0.99
+```
+
+Type I error rate estimation of LatVarfunc
+*Takes a long time to run this, only ran it for sample sizes of 100 and 200*
+```{r}
+set.seed(24253739)
+type_1_error_func <- function(n) {
+  power_result <- matrix(NA, nrow = 100, ncol = 3) 
+  i <- 1
+  repeat {
+
+    sampled_data <- inverted_data[sample(nrow(inverted_data), n, replace = TRUE), ]
+    sampled_data$binary_group <- sample(rep(c(0, 1), each = n / 2))
+
+    inverted_eta <- c(-5.1, -5.1)
+
+    LatVarfunc_result1<- LatVarfunc(sampled_data, inverted_eta)
+
+    power_result[i, ] <- LatVarfunc_result1[1:3]
+    i<- i +1
+    if(i > 100) break}
+  return(sum(power_result[,1] > 0 | power_result[,3] < 0)/ 100)
+}
+sample_vector <- 200
+type_1_error <- sapply(sample_vector, function(n_value) { type_1_error_func(n_value)})
+
+type_1_error_rate_100 <- 0.49
+type_1_error_rate_200 <- 0.57
+```
+
+P-value Estimation of LatVarFunc
+```{r}
+rd <- -0.043
+LCI <- -0.0488
+UCI <- -0.0372
+
+se <- (rd-LCI)/1.96
+to_get_zero <- rd/se
+p_value_latvar <-2*(1-pnorm(abs(to_get_zero)))
 ```
@@
 ```
+
+## References
+ McMenamin, Martina, Jessica K Barrett, Anna Berglind, and James Ms Wason. 2021. “Employing a Latent
+ Variable Framework to Improve Efficiency in Composite Endpoint Analysis.” Statistical Methods in
+ Medical Research 30 (3): 702–16. https://doi.org/10.1177/0962280220970986.

### 2025-06-16 Code Repository.Rmd
updates

@@
 ```{r}
-ggplot(data = null_normal_distribution.df, aes(x =null_normal_distribution )) + geom_histogram(bins =250, colour = "skyblue", fill = "skyblue") + labs(title = "Normal Null Distribution", x = "", y = "Frequency") +
-  theme_minimal() +   geom_vline(xintercept = quantile(null_normal_distribution, probs = c(0.025, 0.975)), color = "black", linetype = "dashed", linewidth = 0.75) + 
-  geom_vline( xintercept = study_summed_composite, color = "gold", size = 1 ) + 
-  annotate("text", x = study_summed_composite, y = 10000, label = "EMERGE Composite", color = "gold", angle = 90, vjust = -0.5)
+ggplot(data = null_normal_distribution.df, aes(x =null_normal_distribution )) + geom_histogram(bins =250, colour = "darksalmon", fill = "darksalmon") + labs(x = "Summed Composite Test Statistic (SC)", y = "Frequency") +
+  theme_minimal() +   geom_vline(xintercept = quantile(null_normal_distribution.df$null_normal_distribution, probs = c(0.025, 0.975)), color = "red", linetype = "dashed", linewidth = 0.75) + 
+  annotate("text", x = quantile(null_normal_distribution.df$null_normal_distribution, 0.025), y = 12000, label = "2.5th Percentile", color = "red", angle = 90, vjust = -0.5, size = 5)+
+  annotate("text", x = quantile(null_normal_distribution.df$null_normal_distribution, 0.975), y = 12000, label = "97.5th Percentile", color = "red", angle = 90, vjust = -0.5, size = 5)+
+  geom_vline( xintercept = study_summed_composite, color = "steelblue", size = 1 ) + 
+  annotate("text", x = study_summed_composite, y = 10000, label = "EMERGE (SC = -9.1)", color = "steelblue", angle = 90, vjust = -0.5)
 ```
@@
 ```{r}
-ggplot(data = theoretical_null_squared, aes(x =chi_sq )) + geom_histogram(bins =250, colour = "lightsalmon", fill = "lightsalmon") + labs(title = "", x = "", y = "") +
+ggplot(data = theoretical_null_squared, aes(x =chi_sq )) + geom_histogram(bins =250, colour = "lightsalmon", fill = "lightsalmon") + labs(x = "Squared Composite Test Statistic (QC)", y = "Frequency") +
   theme_minimal() + 
   geom_vline(xintercept = quantile(theoretical_null_squared$chi_sq, probs = c(0.95)), color = "red", linetype = "dashed", linewidth = 0.75) +
-  annotate("text", x = 8.2, y = 25000, label = "95th Percentile", color = "red", angle = 90, vjust = -0.5, size = 5) +
+  annotate("text", x = quantile(theoretical_null_squared$chi_sq, 0.95), y = 25000, label = "95th Percentile", color = "red", angle = 90, vjust = -0.5, size = 5) +
   theme(panel.grid.minor = element_blank())+
   geom_vline( xintercept = study_squared_composite, color = "steelblue", size = 1 ) + 
-  annotate("text", x = study_squared_composite, y = 25000, label = "Composite", color = "steelblue", angle = 90, vjust = -0.5, size = 5)
-
+  annotate("text", x = study_squared_composite, y = 25000, label = "Composite (QC = 29.2)", color = "steelblue", angle = 90, vjust = -0.5, size = 5)
 ```
@@
 inverted_eta <- c(-5.1, -5.1) ## these are the "cut-off" points for the continuous variables
-results <- LatVarfunc(inverted_data, inverted_eta)
-names(results) <- c("logOR_LCI", "logOR", "logOR_UCI")
+results_lat_var <- LatVarfunc(inverted_data, inverted_eta)
+names(results_lat_var) <- c("logOR_LCI", "logOR", "logOR_UCI", "logRR_LCI", "logRR", "logRR_UCI", "RD_LCI", "RD", "RD_UCI",
+  "Prob_Treat", "Prob_Control", "alpha0", "alpha1", "beta0", "beta1", "psi0", "psi1", "log_sig1", "log_sig2",
+  "logit_rho12", "logit_rho13", "logit_rho23", "se_alpha0", "se_alpha1", "se_beta0", "se_beta1", "se_psi0", "se_psi1",
+  "se_log_sig1", "se_log_sig2", "se_logit_rho12", "se_logit_rho13", "se_logit_rho23")
+results_lat_var_df <- as.data.frame(t(results_lat_var))
+
+log_odds_ratio    <- round(as.numeric(results_lat_var_df[1, 1:3]), 3)
+log_risk_ratio    <- round(as.numeric(results_lat_var_df[1, 4:6]), 3)
+risk_difference   <- round(as.numeric(results_lat_var_df[1, 7:9]), 3)
+
+lat_var_results_table <- data.frame(Measure = c("Log Odds Ratio", "Log Risk Ratio", "Risk Difference"),
+  "Lower Confidence Interval" = c(log_odds_ratio[1], log_risk_ratio[1], risk_difference[1]),
+  `Point Estimate` = c(log_odds_ratio[2], log_risk_ratio[2], risk_difference[2]),
+  `Upper Confidence Interval` = c(log_odds_ratio[3], log_risk_ratio[3], risk_difference[3])
+)
+write.csv(lat_var_results_table, "Latent variable results table.csv", row.names = FALSE)
 ```
@@
 set.seed(24253739)
-power_function_null_dist <- function(n) {
+lat_var_power_function<- function(n) {
   power_result <- matrix(NA, nrow = 100, ncol = 3) 
@@
 }
-sample_vector <- 100
-##null_dist_summed_TS_results <- sapply(sample_vector, function(n_value) {power_function_null_dist(n_value)})
-power_100_latvar<-0.99
-power_150_latvar <- 0.99
+sample_vector <- 200
+lat_var_power_results <- sapply(sample_vector, function(n_value) {lat_var_power_function(n_value)})
+power_100_latvar<-0.83
+power_200_latvar <- 0.97
 ```
@@
 })
-unified_power_results_df <-unified_power_results
 unified_power_results_df <- as.data.frame(t(unified_power_results)) 

### 2025-06-17 Code Repository.Rmd
Minor tweaks

@@
 ## Latent Variable Analysis McMenamin et al.
-The following code is from McMenamin et al. 2021 which has been modified to remove the code for baseline variables of the continuous outcomes.
+The following code is from McMenamin et al. 2021. It has been modified to remove the need for a baseline variable for the continuous variables.
 ```{r}
@@
 Column 5 is Insulin (binary 0, 1)
+
 Make sure to remove ## for the code to run
@@
 inverted_data$gw38 <- inverted_data$gw38 * -1
+inverted_data$baseline1 <- 0
+inverted_data$baseline1[2] <-0.001
+inverted_data$baseline2 <- 0
+inverted_data$baseline2[4] <- 0.001
+
 inverted_eta <- c(-5.1, -5.1) ## these are the "cut-off" points for the continuous variables
@@
 Power estimation of Latent variable analysis
-*Note that this takes a long time to run, and as such we only ran it for two sample sizes, 100 and 150*
+*Note that this takes a long time to run, and as such we only ran it for two sample sizes, 100 and 200*
+Remove the double hashes to execute the power function
 ```{r}
@@
 sample_vector <- 200
-lat_var_power_results <- sapply(sample_vector, function(n_value) {lat_var_power_function(n_value)})
+##lat_var_power_results <- sapply(sample_vector, function(n_value) {lat_var_power_function(n_value)})
 power_100_latvar<-0.83
@@
 *Takes a long time to run this, only ran it for sample sizes of 100 and 200*
+Remove the double hashes to execute the type I error code
 ```{r}
@@
 sample_vector <- 200
-type_1_error <- sapply(sample_vector, function(n_value) { type_1_error_func(n_value)})
+##type_1_error <- sapply(sample_vector, function(n_value) { type_1_error_func(n_value)})
 
-type_1_error_rate_100 <- 0.49
-type_1_error_rate_200 <- 0.57
+LVF_type_1_error_rate_100 <- 0.49
+LVF_type_1_error_rate_200 <- 0.57
 ```
@@
 ```{r}
-power_100_latvar<-0.99
-power_150_latvar <- 0.99
+power_100_latvar<-0.83
+power_200_latvar <- 0.97
 
-LVF_power_df <- data.frame( power = c( 0.99, 0.99),
-                                   sample_size = c(100, 150),
+LVF_power_df <- data.frame( power = c( 0.83, 0.97),
+                                   sample_size = c(100, 200),
                                    label = "Latent Variable")
@@
 
-LVF_type_1_error_rate_100 <- 0.93
-LVF_type_1_error_rate_200 <- 0.96
-LVF_type_1_error_df <- data.frame( error_rate = c( 0.93, 0.96),
+LVF_type_1_error_df <- data.frame( error_rate = c( 0.49, 0.57),
                                    sample_size = c(100, 200),

### 2025-06-25 Code Repository.Rmd
Update Code Repository.Rmd
Wouldn't let me commit remotely

@@
 
+size<-100000
+null_normal_distribution_small <- rnorm(size, mean = 0, sd = sqrt(variance))
+write.csv(null_normal_distribution_small, "Theoretical Summed null distribution small.csv", row.names = F)
+```
+the p-value can be estimated using pnorm() for the asymptotic distribution of the composite (SC)
+```{r}
+p_value_pnorm <- 2*pnorm(study_summed_composite, mean =0, sd = sqrt(variance))
 ```
@@
 
+size <- 100000
+simulated_chi_sq_matrix_small <- matrix(rchisq(size * 3, df = 1), ncol = 3)
+eigen_values_chi_sq_small<- simulated_chi_sq_matrix_small %*%eigen_values  
+write.csv(eigen_values_chi_sq_small, "Theoretical Squared null distribution small.csv", row.names = F)
 ```
@@
 inverted_data$gw38 <- inverted_data$gw38 * -1
-inverted_data$baseline1 <- 0
-inverted_data$baseline1[2] <-0.001
-inverted_data$baseline2 <- 0
-inverted_data$baseline2[4] <- 0.001
 
@@
 power_function <- function(x, y, z, e, n) {
-  summed_TS_power_result <- numeric(1000)  
-  squared_TS_power_result <- numeric(1000)  
-  EMERGE_composite_power <- numeric(1000)
+  summed_TS_power_result <- numeric(10000)  
+  squared_TS_power_result <- numeric(10000)  
+  EMERGE_composite_power <- numeric(10000)
   i <- 1
@@
     i<- i +1
-    if(i > 1000) break}
-  return(as.numeric(c(sum(summed_TS_power_result<quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025)| summed_TS_power_result > quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975))/ 1000,## permutation test summed
-         sum(squared_TS_power_result>quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95))/ 1000, ##permutation test squared
-         sum(squared_TS_power_result > quantile(theoretical_null_squared$chi_sq, 0.95))/ 1000, ##asymptotic chi-square
-         sum((EMERGE_composite_power< 0.05))/ 1000, ## EMERGE
-         sum(summed_TS_power_result<quantile(null_normal_distribution, 0.025)| summed_TS_power_result > quantile(null_normal_distribution, 0.975))/ 1000)))
+    if(i > 10000) break}
+  return(as.numeric(c(sum(summed_TS_power_result<quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025)| summed_TS_power_result > quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975))/ 10000,## permutation test summed
+         sum(squared_TS_power_result>quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95))/ 10000, ##permutation test squared
+         sum(squared_TS_power_result > quantile(theoretical_null_squared$chi_sq, 0.95))/ 10000, ##asymptotic chi-square
+         sum((EMERGE_composite_power< 0.05))/ 10000, ## EMERGE
+         sum(summed_TS_power_result<quantile(null_normal_distribution, 0.025)| summed_TS_power_result > quantile(null_normal_distribution, 0.975))/ 10000)))
 }
@@
 
+##Comparing Asymptotic and permutation null distributions
+```{r}
+asymptotic_null_summed_small <- read.csv("Theoretical Summed null distribution small.csv")
+summed_composite_histogram.df <- data.frame(ts = c(permuted_null_distributions.df$permutation_summed_composite, asymptotic_null_summed_small$V1), group = rep(c("Permuted", "Theoretical"), c(length(permuted_null_distributions.df$permutation_summed_composite), nrow(asymptotic_null_summed_small))))
+
+ggplot(data = summed_composite_histogram.df, aes(x =ts, fill = group )) + geom_histogram(bins =250, alpha = 0.5, position = "identity") + labs(x = "Summed Composite Test Statistic (SC)", y = "Frequency") +
+  theme_minimal() + theme(panel.grid.minor = element_blank()) +
+  scale_fill_manual(values = c("yellow", "skyblue"), name = "Method")
+```
+
+```{r}
+asymptotic_null_sq_small <- read.csv("Theoretical Squared null distribution small.csv")
+
+squared_composite_histogram.df <- data.frame(ts = c(permuted_null_distributions.df$permutation_squared_composite, asymptotic_null_sq_small$V1), group = rep(c("Permuted", "Theoretical"), c(length(permuted_null_distributions.df$permutation_squared_composite), nrow(asymptotic_null_sq_small))))
+
+ggplot(data = squared_composite_histogram.df, aes(x =ts, fill = group )) + geom_histogram(bins =250, alpha = 0.5, position = "identity") + labs(x = "Squared Composite Test Statistic (QC)", y = "Frequency") +
+  theme_minimal() + theme(panel.grid.minor = element_blank()) +
+  scale_fill_manual(values = c("yellow", "skyblue"), name = "Method")
+```
+
 ## References
